---
title: "SPLSDA-which foods/clusters explain phenotypes the best?"
output: html_notebook
---

#libraries
```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("mixOmics")
  

library(tidyverse)
library(mixOmics)
library(plotly)

library(grid)
library(gplots)

# library(pheatmap)
```

# Paths
```{r}
wd <- '../../' 

output_dir = paste0(wd,'results/')
diet.protocols = paste0(output_dir,'diet.protocols/')

data_dir = paste0(wd,'data/')
table_dir = paste0(data_dir, 'tables/')
biom_dir = paste0(data_dir,'silva.feces/')
converter_dir = paste0(output_dir,'converter/')
```

## Create Output Dirs
```{r}
dir.create(output_dir)
dir.create(table_dir)
```

# Data Import
```{r}

keto.bls <- read_tsv(paste0(table_dir, 'keto.bls_04.08.tsv'))

rowSum <- read_tsv(paste0(output_dir,'biomass_30.11.tsv'))

nutr.clust <- read_tsv(paste0(table_dir,'cluster.w.df_07.10.tsv')) 
colnames(nutr.clust)[colnames(nutr.clust) == 'sample'] <- 'file_name'
conv.clust <- read_tsv(paste0(table_dir,'conv.clust.df.weighted.01.10.tsv'))
```
possible to explain data based on food/ nutrient cluster/ conventional cluster intake of converter phenotypes?

sPLS DA with weighted per sample input food/ nutrient cluster/ conventional cluster   and belonging to phenotype

heatmap for intuitive consumtion overview

#food
food distribution explainable through Phenotype?
```{r}
pheno <- dplyr::select(keto.bls,starts_with(c('file_name','Phenotype'))) %>% distinct(.)

pheno$Phenotype[is.na(pheno$Phenotype)] <- 0
pheno$Phenotype[pheno$Phenotype == 0 & pheno$file_name == '34WF.U2'] <- 'not classified'
pheno$Phenotype[pheno$Phenotype == 0 & pheno$file_name == '49RJ.U2'] <- 'low converter'

X.metadata <- left_join(rowSum,pheno,by = 'file_name')

X <- X.metadata %>% dplyr::select(.,everything(),-starts_with(c('subgroup','Phenotype','row'))) %>% 
  column_to_rownames(.,var = 'file_name')

Y = X.metadata$Phenotype %>% as.factor(.)

```

## sparse PLS Discriminant Analysis

### Initial sPLS-DA
```{r}
pseudo = 1e-6

data.plsda = mixOmics::plsda(X = X + pseudo,Y = Y, ncomp = 3 , logratio = 'CLR', scale = T)
data.perf.plsda = mixOmics::perf(data.plsda, validation = 'Mfold', folds = 5,
                    progressBar = FALSE, nrepeat = 10)

plot(data.perf.plsda, sd=TRUE,legend.position = 'horizontal')
mixOmics::plotIndiv(data.plsda , comp = c(1,2),
          group = Y, 
          ind.names = T,
          ellipse = T,
          legend = TRUE, title = 'PLSDA comp 1 - 2')
```

### Cross-Validate features
5x10 CV
```{r}
set.seed(100)
data.tune.splsda = mixOmics::tune.splsda(X + pseudo, 
                          Y = Y, 
                          ncomp = 3, 
                          multilevel = NULL, 
                          logratio = 'CLR',
                          validation = c('Mfold'), 
                          folds = 5, 
                          dist = 'max.dist', 
                          nrepeat = 10,
                          progressBar = FALSE)

plot(data.tune.splsda)
select.keepX = data.tune.splsda$choice.keepX[1:2]
select.keepX # c(15, 5)
```

### Apply CV features
```{r}
data.splsda = mixOmics::splsda(X = X + pseudo,  Y = Y, 
                               ncomp = 3, 
                               keepX = select.keepX, 
                               logratio= "CLR",
                               near.zero.var = T,
)
data.perf.splsda = mixOmics::perf(data.splsda, validation = 'Mfold', folds = 5, 
                   progressBar = FALSE, nrepeat = 10, dist = 'max.dist')

data.perf.splsda$error.rate

mixOmics::plotIndiv(data.splsda , comp = c(1,2),
          group = Y, ind.names = T, 
          ellipse = TRUE, legend = TRUE, title = 'PLSDA comp 1 - 2')
```

Plot loadings 
```{r}
# COMP 1
pL.pc1 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 1',
                       comp = 1, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .4, size.legend = .75)
                       # legend.color = colors.discrete[c(4, 1)])


# COMP 2
# pL.pc2 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 2',
#                        comp = 2, method = 'median', contrib = 'max', 
#                        size.title = rel(1), border = T, 
#                        size.name = .35, size.legend = .3)
#                        # legend.color = colors.discrete[c(4, 1)])

```

#heatmap
weighted food per samples as input
```{r}
X.metadata <- X.metadata %>% dplyr::select(.,everything(),-starts_with('file_name')) 

matrix <- X.metadata %>% group_by(Phenotype) %>%
  summarise_all(.funs = sum) %>%
  pivot_longer(-Phenotype) %>%
  pivot_wider(names_from = Phenotype,values_from=value) %>% 
  column_to_rownames(.,'name') %>% data.matrix(.)

matrix.z <- scale(matrix)
                                                                    stats::heatmap(matrix.z,cexRow = .7,cexCol = .7,main = 'heatmap phenotypes')
                                                                        
#per sample
                                                                        
df <- left_join(rowSum,pheno,by = 'file_name')

df$sample_pheno <- paste(df$file_name,df$Phenotype)
df <- df %>% dplyr::select(.,everything(),-starts_with(c('subgroup','Phenotype','file_name','row')))

matrix.pheno <- df %>% group_by(sample_pheno) %>%
  summarise_all(.funs = sum) %>%
  pivot_longer(-sample_pheno) %>%
  pivot_wider(names_from = sample_pheno,values_from=value) %>% 
  column_to_rownames(.,'name') %>% data.matrix(.)

stats::heatmap(matrix.pheno,cexRow = .7,cexCol = .7,main = 'heatmap phenotypes')                                                                        
```
###export
```{r}
pdf(paste0(converter_dir,'foods.sPLS-DA.pdf'))

mixOmics::plotIndiv(data.splsda , comp = c(1,2),
          group = Y, ind.names = T, 
          ellipse = TRUE, legend = TRUE, title = 'PLSDA foods-phenotypes')
# COMP 1
pL.pc1 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 1',
                       comp = 1, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .7)
                       # legend.color = colors.discrete[c(4, 1)])



stats::heatmap(matrix.z,cexRow = .7,cexCol = .7,main = 'heatmap phenotypes',scale = 'none')

heatmap.2(matrix.z,cexRow = .85,cexCol = .95,trace = 'none',notecol = 'black',srtCol = 30)

dev.off()

```


#nutrient clusters
weighted nutrient cluster as input for sPLS-DA
```{r}

pheno <- dplyr::select(keto.bls,starts_with(c('file_name','Phenotype'))) %>% distinct(.)

pheno$Phenotype[is.na(pheno$Phenotype)] <- 0
pheno$Phenotype[pheno$Phenotype == 0 & pheno$file_name == '34WF.U2'] <- 'not classified'
pheno$Phenotype[pheno$Phenotype == 0 & pheno$file_name == '49RJ.U2'] <- 'low converter'

X.metadata <- left_join(rowSum,pheno,by = 'file_name')

X <- X.metadata %>% dplyr::select(.,everything(),-starts_with(c('subgroup','Phenotype','row'))) %>% 
  column_to_rownames(.,var = 'file_name')

Y = X.metadata$Phenotype %>% as.factor(.)



nutr.clust <- read_tsv(paste0(table_dir,'cluster.w.df_07.10.tsv')) 
colnames(nutr.clust)[colnames(nutr.clust) == 'sample'] <- 'file_name'

pheno.X <- dplyr::select(X.metadata,starts_with(c('Phenotype','file_name')))
# nutr.clust$file_name <- row.names(nutr.clust)

# pheno.X <- pheno.X[grepl('U1',pheno.X$file_name),]
pheno.X <- pheno.X[grepl('U2',pheno.X$file_name),]

nutr.clust <- merge(nutr.clust,pheno,by = 'file_name') %>% column_to_rownames(.,'file_name') %>% dplyr::select(.,everything(),-ends_with('type'))

# nutr.clust <- nutr.clust[grepl('U1',row.names(nutr.clust)),]
nutr.clust <- nutr.clust[grepl('U2',row.names(nutr.clust)),]

Y =pheno.X$Phenotype %>% as.factor(.)

```

## sparse PLS Discriminant Analysis

### Initial sPLS-DA
```{r}
pseudo = 1e-6

#standard Partial Least Squares regression to classify samples

data.plsda = mixOmics::plsda(X = nutr.clust + pseudo,Y = Y, ncomp = 2 , logratio = 'CLR', scale = T)
data.perf.plsda = mixOmics::perf(data.plsda, validation = 'Mfold', folds = 5,
                    progressBar = FALSE, nrepeat = 10,auc = TRUE)

plot(data.perf.plsda, sd=TRUE,legend.position = 'horizontal')
mixOmics::plotIndiv(data.plsda , comp = c(1,2),
          group = Y, 
          ind.names = T,
          ellipse = T,
          legend = TRUE, title = 'PLSDA phenotypes')
```

### Apply Cross Validation features
prediction whether sample belongs to sg1 or sg2
Cross Validation tries how stable prediction is
```{r}
data.splsda = mixOmics::splsda(X = nutr.clust + pseudo,  Y = Y, 
                               ncomp = 2, 
                               # keepX = select.keepX, 
                               logratio= "CLR",
                               near.zero.var = T,
)
data.perf.splsda = mixOmics::perf(data.splsda, validation = 'Mfold', folds = 5, 
                   progressBar = FALSE, nrepeat = 10, dist = 'max.dist')

data.perf.splsda$error.rate

h <- mixOmics::plotIndiv(data.splsda , comp = c(1,2),
          group = Y, ind.names = T, 
          ellipse = TRUE, legend = TRUE, title = 'PLSDA phenotypes')


```


Plot loadings 
```{r}
# COMP 1
pL.pc1 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 1',
                       comp = 1, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .75)
                       # legend.color = colors.discrete[c(4, 1)])


# COMP 2
pL.pc2 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 2',
                       comp = 2, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .75)
                       # legend.color = colors.discrete[c(4, 1)])


plotIndiv(data.splsda)
```

#heatmap
weighted nutrient clusters as input
```{r}
nutr.clust <- read_tsv(paste0(table_dir,'cluster.w.df_07.10.tsv'))
colnames(nutr.clust)[colnames(nutr.clust) == 'sample'] <- 'file_name'


nutr.clust <- merge(nutr.clust,pheno,by = 'file_name') %>% column_to_rownames(.,'file_name') 

# nutr.clust <- nutr.clust[grepl('U1',row.names(nutr.clust)),]
nutr.clust <- nutr.clust[grepl('U2',row.names(nutr.clust)),]

matrix.nutr <- nutr.clust %>% group_by(Phenotype) %>%
  summarise_all(.funs = sum) %>%
  pivot_longer(-Phenotype) %>%
  pivot_wider(names_from = Phenotype,values_from=value) %>% 
  column_to_rownames(.,'name') %>% data.matrix(.)

matrix.nutr.z <- scale(matrix.nutr)
                                                                  
heatmap.2(matrix.nutr.z,cexRow = .85,cexCol = .95,trace = 'none',notecol = 'black',srtCol = 30)

l <- stats::heatmap(matrix.nutr.z,cexRow = .9,cexCol = .9,Colv = NA,Rowv = NA)
                                                                        
#per sample
                                                                        
                                                                        
df <- nutr.clust
df$file_name <- row.names(df)

df$sample_pheno <- paste(df$file_name,df$Phenotype)
df <- df %>% dplyr::select(.,everything(),-starts_with(c('subgroup','Phenotype','file_name')))

matrix.pheno <- df %>% group_by(sample_pheno) %>%
  summarise_all(.funs = sum) %>%
  pivot_longer(-sample_pheno) %>%
  pivot_wider(names_from = sample_pheno,values_from=value) %>% 
  column_to_rownames(.,'name') %>% data.matrix(.)


k <- stats::heatmap(matrix.pheno,cexRow = .7,cexCol = .7,verbose = TRUE) 

# mp1 <- matrix.pheno[,grepl(pattern = 'U1',colnames(matrix.pheno))]
# mp2 <- matrix.pheno[,grepl(pattern = 'U2',colnames(matrix.pheno))]
# mp <- cbind(mp1,mp2)
# stats::heatmap(mp,cexRow = .7,cexCol = .7,main = 'heatmap phenotypes') 
```
###export
```{r}
pdf(paste0(converter_dir,'U2.nutclust.sPLS-DA.pdf'))

h

# COMP 1
pL.pc1 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA nutrient clusters',
                       comp = 1, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .7)
                       # legend.color = colors.discrete[c(4, 1)])


# COMP 2
pL.pc2 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 2',
                       comp = 2, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .7)
                       # legend.color = colors.discrete[c(4, 1)])

stats::heatmap(matrix.nutr.z,cexRow = .9,cexCol = .9,Colv = NA,Rowv = NA,scale = 'none')

stats::heatmap(matrix.pheno,cexRow = .7,cexCol = .7,scale = c('row')) 

heatmap.2(matrix.nutr.z,cexRow = .85,cexCol = .75,trace = 'none',notecol = 'black',srtCol = 30)

dev.off()

```


#conv cluster
weighted conventional clusters as sPLS DA input
```{r}
pheno <- dplyr::select(keto.bls,starts_with(c('file_name','Phenotype'))) %>% distinct(.)

pheno$Phenotype[is.na(pheno$Phenotype)] <- 0
pheno$Phenotype[pheno$Phenotype == 0 & pheno$file_name == '34WF.U2'] <- 'not classified'
pheno$Phenotype[pheno$Phenotype == 0 & pheno$file_name == '49RJ.U2'] <- 'low converter'

X.metadata <- left_join(rowSum,pheno,by = 'file_name')

X <- X.metadata %>% dplyr::select(.,everything(),-starts_with(c('subgroup','Phenotype','row'))) %>% 
  column_to_rownames(.,var = 'file_name')

Y = X.metadata$Phenotype %>% as.factor(.)

conv.clust <- read_tsv(paste0(table_dir,'conv.clust.df.weighted.01.10.tsv'))


pheno.X <- dplyr::select(X.metadata,starts_with(c('Phenotype','file_name')))
# nutr.clust$file_name <- row.names(nutr.clust)

# pheno.X <- pheno.X[grepl('U1',pheno.X$file_name),]
# pheno.X <- pheno.X[grepl('U2',pheno.X$file_name),]

conv.clust <- merge(conv.clust,pheno,by = 'file_name') %>% column_to_rownames(.,'file_name') %>% dplyr::select(.,everything(),-ends_with('type'))

# conv.clust <- conv.clust[grepl('U1',row.names(conv.clust)),]
# conv.clust <- conv.clust[grepl('U2',row.names(conv.clust)),]

Y =pheno.X$Phenotype %>% as.factor(.)

```

## sparse PLS Discriminant Analysis

### Initial sPLS-DA
```{r}
pseudo = 1e-6

#standard Partial Least Squares regression to classify samples

data.plsda = mixOmics::plsda(X = conv.clust + pseudo,Y = Y, ncomp = 2 , logratio = 'CLR', scale = T)
data.perf.plsda = mixOmics::perf(data.plsda, validation = 'Mfold', folds = 5,
                    progressBar = FALSE, nrepeat = 10)

plot(data.perf.plsda, sd=TRUE)
mixOmics::plotIndiv(data.plsda , comp = c(1,2),
          group = Y, 
          ind.names = T,
          ellipse = T,
          legend = TRUE, title = 'PLSDA phenotypes')
```

### Apply Cross Validation features
prediction whether sample belongs to sg1 or sg2
Cross Validation tries how stable prediction is
```{r}
data.splsda = mixOmics::splsda(X = conv.clust + pseudo,  Y = Y, 
                               ncomp = 2, 
                               # keepX = select.keepX, 
                               logratio= "CLR",
                               near.zero.var = T,
)
data.perf.splsda = mixOmics::perf(data.splsda, validation = 'Mfold', folds = 5, 
                   progressBar = FALSE, nrepeat = 10, dist = 'max.dist')

data.perf.splsda$error.rate

h <- mixOmics::plotIndiv(data.splsda , comp = c(1,2),
          group = Y, ind.names = T, 
          ellipse = TRUE, legend = TRUE, title = 'PLSDA phenotypes')


```


Plot loadings 
```{r}
# COMP 1
pL.pc1 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 1',
                       comp = 1, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .45)
                       # legend.color = colors.discrete[c(4, 1)])


# COMP 2
pL.pc2 <- mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 2',
                       comp = 2, method = 'median', contrib = 'max', 
                       size.title = rel(1), border = T, 
                       size.name = .75, size.legend = .75)
                       # legend.color = colors.discrete[c(4, 1)])

```

#heatmap
weighted conventional clusters as input
```{r}
conv.clust <- read_tsv(paste0(table_dir,'conv.clust.df.weighted.01.10.tsv'))

conv.clust <- merge(conv.clust,pheno,by = 'file_name') %>% column_to_rownames(.,'file_name') 

# conv.clust <- conv.clust[grepl('U2',row.names(conv.clust)),]

matrix.conv <- conv.clust %>% group_by(Phenotype) %>%
  summarise_all(.funs = sum) %>%
  pivot_longer(-Phenotype) %>%
  pivot_wider(names_from = Phenotype,values_from=value) %>% 
  column_to_rownames(.,'name') %>% data.matrix(.)

matrix.conv.z <- scale(matrix.conv)                                                                 

l <- stats::heatmap(matrix.conv.z,cexRow = .8,cexCol = .9,Rowv = NA,Colv = NA)
                                                                    heatmap.2(matrix.conv.z,cexRow = .85,cexCol = .95,trace = 'none' ,notecol = 'black',srtCol = 30)  
#per sample
                                                                        
                                                                        
df <- conv.clust
df$file_name <- row.names(df)

df$sample_pheno <- paste(df$file_name,df$Phenotype)
df <- df %>% dplyr::select(.,everything(),-starts_with(c('subgroup','Phenotype','file_name')))

matrix.pheno <- df %>% group_by(sample_pheno) %>%
  summarise_all(.funs = sum) %>%
  pivot_longer(-sample_pheno) %>%
  pivot_wider(names_from = sample_pheno,values_from=value) %>% 
  column_to_rownames(.,'name') %>% data.matrix(.)


k <- stats::heatmap(matrix.pheno,cexRow = .7,cexCol = .7) 

# mp1 <- matrix.pheno[,grepl(pattern = 'U1',colnames(matrix.pheno))]
# mp2 <- matrix.pheno[,grepl(pattern = 'U2',colnames(matrix.pheno))]
# mp <- cbind(mp1,mp2)
# stats::heatmap(mp,cexRow = .7,cexCol = .7,main = 'heatmap phenotypes') 
```
###export
```{r}
pdf(paste0(converter_dir,'convclust.sPLS-DA.pdf'))

h

# COMP 1
mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA conventional clusters',
                       comp = 1, method = 'median', contrib = 'max',
                       size.title = rel(1), border = T,
                       size.name = .75, legend = FALSE)
                       # legend.color = colors.discrete[c(4, 1)])


# COMP 2
mixOmics::plotLoadings(data.splsda, title = 'sPLS-DA PC 2',
                       comp = 2, method = 'median', contrib = 'max',
                       size.title = rel(1), border = T,
                       size.name = .75, legend = FALSE)
                       # legend.color = colors.discrete[c(4, 1)])


heatmap.2(matrix.conv.z,cexRow = .85,cexCol = .95,trace = 'none',notecol = 'black',srtCol = 30) 

stats::heatmap(matrix.conv.z,cexRow = .8,cexCol = .9,Rowv = NA,Colv = NA,scale = 'none')  
#per sample
# COMP 1


dev.off()

```
high and low converter U1 and U1 getrennt
```{r}

```

