---
title: "BLS beta-diversity"
output: html_notebook
---

# Libraries
```{r}
library(tidyverse)
library(Rtsne)
library(FactoMineR)

library(rgr)
library(colorspace)
library(dendextend)

library(factoextra)
library(ape)
library(phytools)
library(ggpubr)
```

# Paths
```{r}
output_dir = '../results/'
output_cluster = paste0(output_dir,'cluster/')
data_dir = '../data/'
table_dir = paste0(data_dir, 'tables/')

```

# Data Import
```{r}
bls.scale <- read_tsv(paste0(table_dir, 'BLS.scale.15Jun2020.tsv'))
bls.properties <- read_tsv(paste0(table_dir, 'BLS.properties.15Jun2020.tsv'))

keto.bls <- read_tsv(paste0(table_dir, 'keto.bls_04.08.tsv'))
# subclusters <- read_tsv(paste0(table_dir, 'subclusters_06.10.tsv'))

# nut.clust <- read_tsv(paste0(table_dir, 'food.conv.nutrient.tsv'))
keto.bls.merged <- read_tsv(paste0(table_dir, 'keto.bls.merged.tsv'))%>%
  dplyr::select(.,starts_with('item'))
```

PCA of BLS foods

nutrient clustering based on PCA (hclust cut at height 95%)

PCA colored by conventional clusters 
PCA colored by nutrient clusters

tSNE of BLS foods

# Dimensionality Reduction

Data
```{r}
keep <- !bls.scale$item_cat %in% c('Gewürze, Würzmittel, Hilfsstoffe')
#%in% c() #'Menükomponenten','Gewürze, Würzmittel, Hilfsstoffe' 
metadata <- bls.scale[keep, c('item_id','item_cat','item_subcat','item_descr')]
df <- bls.scale[keep,] %>%
  dplyr::select(everything(), -starts_with('item'), item_id) %>%
  mutate_at(.vars = vars(everything(), -item_id), 
            .funs = funs(replace_na(., 0))) %>% 
  column_to_rownames('item_id') %>% 
  as.matrix()

```


## PCA

### gelman normalization

Transform
```{r}
scale_gelman <- function(x) {
  return((x - mean(x, na.rm = T)) / (2 * sd(x, na.rm = T)))
}
df.scale_gelman <- apply(df, 2, function(x) scale_gelman(x))
```

```{r}
df.scale_gelman.pca_data <- PCA(X = df.scale_gelman, graph = FALSE, scale.unit = T)

df.scale_gelman.pca <- df.scale_gelman.pca_data %>% 
  .$ind %>% .$coord %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'Name') 

df.data.scale_gelman.pca <- df.scale_gelman.pca %>% 
  left_join(., metadata, by = c('Name' = 'item_id'))
```

```{r fig.width=10, fig.height=10}
df.data.scale_gelman.pca %>% 
  ggplot(aes(Dim.2, Dim.1)) +
  geom_point(size = 0.5, aes(col = item_cat)) + 
  labs(y = paste0('PC 1 (', round(df.scale_gelman.pca_data$eig[1,'percentage of variance'], 1), '%)'), 
       x = paste0('PC 2 (', round(df.scale_gelman.pca_data$eig[2,'percentage of variance'], 1), '%)')) +
  theme(plot.title = element_text(hjust = 0.5),
        aspect.ratio = 1,
        # legend.position = 'none'
        ) + 
  guides(fill = guide_legend(ncol = 1))
```

### clr transform

```{r}
pseudo = 1e-15 # 1e-9
df.clr <- rgr::clr(df + pseudo, ifclose = FALSE, ifwarn = TRUE)
df.clr.unit <- t(scale(t(df.clr), center = F, scale = T))
```

```{r}
X = cbind(metadata$item_cat, as.data.frame(df.clr))
df.clr.pca_data <- PCA(X = df.clr, graph = FALSE, scale.unit = T)

df.clr.pca <- df.clr.pca_data %>% 
  .$ind %>% .$coord %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'Name') 

df.data.clr.pca <- df.clr.pca %>% 
  left_join(., metadata, by = c('Name' = 'item_id'))

write_tsv(df.data.clr.pca ,paste0(table_dir,'df.data.clr.pca.tsv'))


```

Plot
```{r fig.width=15, fig.height=10}
df.data.clr.pca %>% 
  ggplot(aes(Dim.1, Dim.2, labels = item_descr)) +
  geom_point(size = 0.5, aes(col = item_cat)) + 
  labs(y = paste0('PC 1 (', round(df.clr.pca_data$eig[1,'percentage of variance'], 1), '%)'), 
       x = paste0('PC 2 (', round(df.clr.pca_data$eig[2,'percentage of variance'], 1), '%)')) +
  theme(plot.title = element_text(hjust = 0.5),
        aspect.ratio = 1, 
        ) + 
  guides(fill = guide_legend(ncol = 1))
```


Loop through Property Categories
```{r fig.width=15, fig.height=10}
pexpand <- rep(0.1, 2)
prop_cat <- 
  tibble(Property = colnames(df.clr)) %>% 
  left_join(., bls.properties) %>% 
  mutate(Category = replace_na(Category, 'Zusammensetzung')) %>% 
  pull(Category)

# k <- dplyr::select(keto.bls,starts_with = c('item_id','item_prodi')) %>% distinct()

p <-
  factoextra::fviz_pca_biplot(X = df.clr.pca_data, 
                            # samples
                            fill.ind = metadata$item_cat, col.ind = 'black',
                            pointshape = 21, pointsize = 1.5,
                            geom.ind = 'point', repel = T, geom.var = FALSE) +
                            
                            # variables
                            # col.var = prop_cat,
                            # title = 'Overview of all Features',
                            # label = F) + 
  theme(aspect.ratio = 1) + 
  scale_x_continuous(expand = pexpand) + 
  scale_y_continuous(expand = pexpand)




                            
                            
for (cat in unique(bls.properties$Category)) {
  svars = bls.properties %>% filter(Category == cat) %>% pull(Property)
  p <- factoextra::fviz_pca_biplot(X = df.clr.pca_data, 
                            # samples
                            fill.ind = metadata$item_cat, col.ind = 'black',
                            pointshape = 21, pointsize = 1.5,
                            geom.ind = 'point', repel = T, 
                            
                            # variables
                            col.var = 'black',
                            select.var = list('name' = svars), 
                            title = paste0('Feature: ', cat)) + 
  theme(aspect.ratio = 1) + 
  scale_x_continuous(expand = pexpand) + 
  scale_y_continuous(expand = pexpand)
  
  fname = paste0(output_dir, 'BLS.', cat, '.pdf')
  print(fname)
  print(p)
  # ggsave(p, filename = fname, device = 'pdf', width = 15, height = 10, dpi = 300)
}

#extract PCs

pca <- get_pca(df.clr.pca_data,'ind')
# plot(pca$coord)
pca.coord.bls <- as.data.frame(pca$coord)



```


##PC clustering
nutrient food clustering based on PCS of BLS foods
PC (variables) are the nutrients

hclust based on first three axis, cut at 97% similarity
```{r}
##extract PC and calculate distance

keto <- dplyr::select(keto.bls,starts_with(c('item_id','item_prodi'))) %>% distinct(.) 
df <- df.clr.pca_data$ind$coord %>% as.data.frame(.) %>% rownames_to_column(.)
names(df)[names(df) == 'rowname'] <- 'item_id'
coord <- merge(df,keto,all.x = TRUE,by = 'item_id')
coord$item_prodi[is.na(coord$item_prodi)] <- 'nc'
coord <- dplyr::filter(coord, item_prodi != "nc")

#variance axis
varAxis.1 <- var(coord$Dim.1)##30.73313
varAxis.2 <- var(coord$Dim.2)##25.47235
varAxis.3 <- var(coord$Dim.3)##31.53042

coord <- dplyr::select(coord,everything(),-starts_with(c('item_id','Dim.4','Dim.5', 'Dim.3'))) %>% column_to_rownames(.,'item_prodi')
```


```{r}
#distance,hclust,tree

pc.hclust <- coord %>% dist(.,method = 'euclidean') %>%
  hclust(.,method = 'ward')
dend.nutr <- as.dendrogram(pc.hclust)

pc.tree <- ape::as.phylo(pc.hclust)
# cut = cutree(pc.hclust, k = 10) %>% as.data.frame() %>% rownames_to_column('food') %>% rename(group = '.')
cut = cutree(pc.hclust, h = (max(pc.hclust$height)) * (1 - 0.97)) %>% as.data.frame() %>% rownames_to_column('food') %>% rename(group = '.')

# write_tsv(cut,paste0(table_dir,nutrient.clusters.09.09.tsv',append = FALSE))

# write.tree(pc.tree, file=paste0(table_dir,'tree.keto.bls.19.08.tre',append = FALSE))
```
PCA coloored by nutrient clusteers
```{r}
###color by conv cluster

pc.nut <- coord %>% rownames_to_column('food') %>% 
  full_join(., cut) 
pc.nut$group <- as.character(pc.nut$group)

colnames(metadata)[colnames(metadata) == 'item_descr'] <- 'food'
metadata$food <- tolower(metadata$food)
pc <- left_join(pc.nut,metadata,by = 'food')
pc[is.na(pc)] <- 'nc'

pc.n <- pc[!grepl("nc", pc$item_cat),]

cc <- ggplot(pc.n,aes(Dim.1, Dim.2)) + 
  geom_point(aes(fill = as.factor(item_cat)), shape = 21,size = 1) +
   theme(aspect.ratio = 1,legend.key.width = unit(0.35,'cm'),
         legend.text = element_text(size = 3)) + 
  scale_x_continuous(expand = pexpand) + 
  scale_y_continuous(expand = pexpand) + coord_cartesian(xlim=c(-15,15), ylim=c(-15,15))


# pdf(paste0(output_dir,'pc.conv.pdf',width = 15,height = 15))
cc
# dev.off()

###color by nutrient cluster

pc.n$group[pc.n$group == '1'] <- '1.veg.fruit'
pc.n$group[pc.n$group == '2'] <- '2.carbs.bread'
pc.n$group[pc.n$group == '3'] <- '3.fatty.sweet'
pc.n$group[pc.n$group == '4'] <- '4.carbs.cereal'
pc.n$group[pc.n$group == '5'] <- '5.seed'
pc.n$group[pc.n$group == '6'] <- '6.nut'
pc.n$group[pc.n$group == '7'] <- '7.kohl'
pc.n$group[pc.n$group == '8'] <- '8.fish.sauce'
pc.n$group[pc.n$group == '9'] <- '9.egg'
pc.n$group[pc.n$group == '10'] <- '10.alcohol'
pc.n$group[pc.n$group == '11'] <- '11.wurst'
pc.n$group[pc.n$group == '12'] <- '12.soft.drinks'
pc.n$group[pc.n$group == '13'] <- '13.white.meat'
pc.n$group[pc.n$group == '14'] <- '14.fatty.dairy'
pc.n$group[pc.n$group == '15'] <- '15.cheese'
  
c <- ggplot(pc.n,aes(Dim.1, Dim.2)) + 
  geom_point(aes(fill = as.factor(group)), shape = 21,size = 1) +
   theme(aspect.ratio = 1,legend.key.width = unit(0.35,'cm'),legend.text = element_text(size = 3),plot.title = element_text('nutrient food clusters')) + 
  scale_x_continuous(expand = pexpand) + 
  scale_y_continuous(expand = pexpand)+ coord_cartesian(xlim=c(-15,15), ylim=c(-15,15))



# pdf(paste0(output_dir,'pc.nutrient.pdf',width = 10, height = 10))
c
# dev.off()
pdf(paste0(output_cluster,'pca_clusters.pdf'))
ggarrange(cc,c,ncol = 2,nrow = 1,legend = "none")
cc
c
dev.off()

a <- table(pc.n$group) %>% as.data.frame(.)
b <- table(pc.n$item_cat) %>% as.data.frame(.)

colnames(a) <- c('nutrient cluster','number of foods')
colnames(b) <- c('conventional cluster','number of foods')

pdf(paste0(output_cluster,'nutrient.cluster.count.pdf'))
grid.table(a)
dev.off()

pdf(paste0(output_cluster,'conv.cluster.count.pdf'))
grid.table(b)
dev.off()
```

#conv clustering

PCA BLS food colored by conventional clusters
```{r}
keto.bls <- read_tsv(paste0(table_dir, 'keto.bls_04.08.tsv'))
keto.bls <- keto.bls[!(keto.bls$item_cat == 'Gewürze, Würzmittel, Hilfsstoffe' ),]##| keto.bls$item_cat =='Menükomponenten'
tmp <- unique(subset(keto.bls,select = c(`item_cat`,`item_subcat`,`item_prodi`))) %>% distinct(.)
r <- tmp
colnames(r) <- c('item_cat.select','item_subcat.select','item_prodi.select') 



r <- tidyr::uncount(r, dplyr::n())
# tmp <- tmp[rep(seq_len(nrow(tmp)), nrow(r)), ]
tmp <- purrr::map_dfr(seq_len(1250), ~tmp)##nrow(tmp)/nrow(r)=1044  ##w/o gewürze 1280
tmp <- cbind(tmp,r)

cat.dist <-
  tmp %>% 
  mutate(same_subcat = item_subcat.select == item_subcat,
         same_cat = item_cat.select == item_cat) %>% 
  mutate(total = (as.numeric(!same_cat) + as.numeric(!same_subcat)) / 2) %>% 
  dplyr::select(item_prodi.select, item_prodi, total) %>%
  
  pivot_wider(names_from = 'item_prodi.select', values_from = 'total') %>% 
  column_to_rownames('item_prodi') %>%
  as.dist(.)

cat.dist[is.na(cat.dist)] <- 0

d.cat_data <- hclust(cat.dist, method = 'complete')
dend.cat <- as.dendrogram(d.cat_data)
tree.conv <- ape::as.phylo(d.cat_data,use.labels = keto.bls$item_cat)
```
## t-SNE
```{r}
df.tsne_data <- Rtsne(X = as.matrix(df.clr),
                     check_duplicates = FALSE,
                     pca = T, pca_center = T, pca_scale = T, normalize = T, 
                     dims = 2, perplexity = 35,
                     verbose=TRUE, max_iter=1000)

df.data.tsne <- cbind(metadata, df.tsne_data$Y) %>% 
  rename(Dim.1 = `1`, 
         Dim.2 = `2`) 
```

Plot
```{r fig.width=10, fig.height=10}
df.data.tsne %>% 
  ggplot(aes(Dim.2, Dim.1)) +
  geom_point(size = 0.5, aes(col = item_cat)) + 
  theme(plot.title = element_text(hjust = 0.5),
        aspect.ratio = 1, 
        ) + 
  guides(fill = guide_legend(ncol = 1))
```

### Loop Perplexity
```{r}
loop_tsne <- function(df, label) {
  output <- NULL
  for (perplexity in seq(10, 50, 20)) {
  tmp.tsne_data <- Rtsne(X = as.matrix(df),
                     check_duplicates = FALSE,
                     pca = T, pca_center = T, pca_scale = T, normalize = T, 
                     dims = 2, perplexity = perplexity,
                     verbose=TRUE, max_iter=400)
  tmp.data <- 
    cbind(metadata, tmp.tsne_data$Y) %>% 
    rename(Dim.1 = `1`, 
         Dim.2 = `2`) %>% 
    mutate(perplexity = perplexity)  
  output <- rbind(output, tmp.data)
  }
  return(output %>% mutate(label = label))
}

tsne.clr <- loop_tsne(df.clr, 'clr')
tsne.raw <- loop_tsne(df, 'raw')
tsne.scale_gelman <- loop_tsne(df.scale_gelman, 'gelman')
```



```{r}
df.tsne_loop.all <- 
  bind_rows(.id = 'method',
    df = tsne.raw, 
    df.clr = tsne.clr, 
    df.scale_gelman = tsne.scale_gelman)
```


```{r fig.width=20, fig.height=15}
df.tsne_loop.all %>% 
  filter(!item_cat %in% c('Gewürze, Würzmittel, Hilfsstoffe')) %>%  #'Menükomponenten' ,
  ggplot(aes(Dim.2, Dim.1)) +
  geom_point(size = 1, alpha = 0.5, aes(col = item_cat)) + 
  theme(plot.title = element_text(hjust = 0.5),
        aspect.ratio = 1,
        ) + 
  guides(fill = guide_legend(ncol = 1)) + 
  theme(legend.position = 'none') + 
  facet_grid(rows = vars(method), 
             cols = vars(perplexity), scales = 'free', space = 'free')
```

# hclust

```{r}
df.clr.pca.hclust_data <- hclust(dist(df.data.clr.pca %>% dplyr::select(starts_with('Dim'))), method = 'ward.D')
plot(df.clr.pca.hclust_data)

#how many categories,subcategories are there?

count_cat <- length(unique(bls.gram[["item_cat"]]))##21
count_subcat <- length(unique(bls.gram[["item_subcat"]]))#178

```
