---
title: "animal and plant derived items"
output: html_notebook
---
#libraries
```{r}
library(tidyverse)
library(plotly)
library(ggplot2)
library(ggsignif)
library(ggpubr)

library(grid)
```

# Paths
```{r}

wd <- '../../' 

output_dir = paste0(wd,'results/')
diet.protocols = paste0(output_dir,'diet.protocols/')

data_dir = paste0(wd,'data/')
table_dir = paste0(data_dir, 'tables/')
biom_dir = paste0(data_dir,'silva.feces/')
```

## Create Output Dirs
```{r}
dir.create(wd)
dir.create(output_dir)
dir.create(table_dir)
```

#data import
```{r}
k.bls_prodi <- read_tsv(paste0(table_dir,'k.bls_prodi_13_08.tsv'))

k.bls_prodi$file_name <- gsub(k.bls_prodi$file_name,pattern='u',replacement = 'U') %>% gsub(pattern = '_',replacement = '.')
```


#abundance of animal derived items
manually assigned
```{r}

animal <- k.bls_prodi[k.bls_prodi$item_subcat %in% c('Likör (durchschn. 32 %)','Brühwürfel, Fleischextrakt','Kuchen und Torten aus besonderen Teigen','Fertiggerichte auf Fleischwarenbasis','Innereien von Haustieren','Fischerzeugnisse','Krebstiere, Muscheltiere','Schweinefleischstücke','Kalbfleischstücke','Barschartige Fische','Obstkuchen','Rindfleischschnitte','Dorschartige Fische, Kabeljau, Schellfisch, Seelachs, Goldbarsch','Torten','Fleischgerichte v. Wild + Geflügel','Fleischgerichte vom Hammel+Lamm','Schweinefleischschnitte','Fleischwarenerzeugnisse','Fischgerichte','Speiseeis','Milcherzeugnisse und Käseerzeugnisse','Ölsoßen, Mayonnaisen, Fettzubereitungen','Heringsfische, Makrelen, Thunfische','Obsttorten','Schnittkäse halbfest','Tierische Fette','Fleischgerichte vom Kalb','Getreidegerichte, Mehlspeisen, Pfannkuchen','Milchmischerzeugnisse','Pökelwaren','Geflügel','Schokoladenwaren, Pralinen','Teigwaren (-arten)','Eier-, Quark- und Käsegerichte','Kochwurst','Lachsfische','Würstchen, Hackfleischgerichte Fleisch,','Kalbfleischschnitte','Fleischgerichte vom Schwein','Frischkäse, Sauermilchkäse, Kochkäse, Schmelzkäse','Brühwurst','Süßspeisen, Dessert','Fertiggerichte auf Fleischbasis','Fleischgerichte vom Rind','Soßen','Hartkäse','Rindfleischstücke','Nudelgerichte, Teigwarengerichte, Pizzen','Marzipan, Lakritze, Krokant, Nougat','Speck, Schinken','Schnittkäse','Weichkäse','Schokolade','Rohwurst','Kuchen','Schnellgerichte, Eis','Eier','Butter','Milch und Sauermilch'),]

animal2 <- k.bls_prodi[k.bls_prodi$item %in% c('mozzarella','french dressing energiereduziert','choco crossies nestle','gemüsequiche (standardrezeptur)','feta','rind hackfleisch roh','blätterteig gebacken','schafskäse','rind/schwein hackfleisch gekocht','schafskäse mind. 50% fett i. tr.','schafskäse','dressing für kartoffelsalat fertigprodukt','schwein hackfleisch gegart','rind/schwein hackfleisch gebraten (zubereitet ohne fett)','wurst','kapern konserve abgetropft','zitronenkuchen (standardrezeptur)','french dressing','buttergemüse (tk), alnatura','	
hacksteak','rinderhackfleisch','rind hackfleisch gekocht (zubereitet mit fett und salz)','waffelkeks'),]

animal <- rbind(animal,animal2)

#merging file_name and day together in a new column sample_day

animal$sample_day <- paste(animal$file_name,animal$day)
animal$source <- rep('animal',nrow(animal))
```

#abundance of plant derived items
manually assigned 
```{r}

plant <- k.bls_prodi[k.bls_prodi$item_subcat %in% c('Hülsenfruchtgerichte','Sojaerzeugnisse','Branntwein aus Wein','Kerne','Vegetarische Erzeugnisse','Fruchtsaftgetränk','Nusserzeugnisse und Ölsamenerzeugnisse','Knäckebrot','Pflanzliche Öle, 30–60 % Linolsäuregehalt','Gewürze','Sojamilch und Sojamilchprodukte','Eiweißmodifizierte Lebensmittel','Getreide','Obsterzeugnisse','Hülsenfrüchte (reife)','Gemüseerzeugnisse','Weißbrot, Feinbrot','Sprossen, Keime','Kaffeeersatz','Reisgerichte','Ölsamen','Fruchtzubereitungen','Spezialbrote/Spezialbrötchen','Broterzeugnis','Pilzerzeugnisse','Teigwaren besonderer Art','Weinähnliche Getränke, Schaumwein (bis 18 %)','Getreideerzeugnisse','Stärkereiche Pflanzenteile','Vegetarische Fertiggerichte','Vollkornbrot','Zitrusfrüchte','Kaffee','Pflanzliche Fette','Vollkornbrötchen (Kleingebäck)','Salate, gegart','Früchtetee, Kräutertee','Spirituosen (min. 32–40 %)','Blattgemüse','Kohlgemüse','Zuchtpilze/Wildpilze','Anderes Schalenobst','Beerenobst','Steinobst','Brötchen, (Kleingebäck)','Tee','Weißwein, Rotwein, Roséwein (durchschn. 10 %)','Kartoffeln','Hülsenfruchtgemüse unreif','Kakao/Kakaogetränk','Kleinteile, Plätzchen, Keks','Kartoffelgerichte','Cocktails','Bier (durchschn. 3,5 %)','Nüsse','Salate, roh','Gemüsebeilagen','Wildfrüchte','Sprossengemüse und Lauchgemüse','Pflanzliche Öle, >60 % Linolsäuregehalt','Brausen/Limonaden','Kartoffelerzeugnisse','Pflanzliche Öle, <30 % Linolsäuregehalt','Spezielle Getreide','Suppen','Brot (Graubrot, Schwarzbrot, Mischbrot)','Ölfrüchte','Fruchtgemüse','Wurzelgemüse und Knollengemüse (außer Kartoffeln)','Salatgemüse','Hilfsmittel, Backtriebe, Geliermittel und Dickungsmittel','Mehl','Zuckerwaren, Bonbons','Würzmittel','	
Margarine','Mineralwasser, Trinkwasser','Südfrüchte','Kernobst','Belegte Brote, Toasts, Frühstückscerealien','Reduktionskost'),]

plant2 <- k.bls_prodi[k.bls_prodi$item %in% c('laugenbrezel (mit salz)','hefe','chia','salzstangen (salzbrezeln) als dauergebäck','gemüsemischung chinesische art gebraten (zubereitet ohne fett)','sojasoße','hefeteig leicht gebacken','gemüsemischung tiefgefroren gegart','gemüsemischung roh','gemüse','müsli-riegel','rhabarber fruchtsaft','hefeaufstrichpaste mit olive','gemüsemischung gegart','gemüsemischung gekocht','gemüsebeilagen (0)','cracker meersalz, alnatura','tuc classic cracker, griesson - de beukelaer','cracker','gemüsemischung chinesische art gegart','sellerieblätter frisch','gemüsemischung chinesische art gebacken','laugenbrezel (sichtbares salz entfernt)','rind/schwein hackfleisch roh','suppengrün tiefgefroren'),]

plant <- rbind(plant,plant2)

plant$sample_day <- paste(plant$file_name,plant$day)
plant$source <- rep('plant',nrow(plant))
```

#density dot plot
x and y axis plant and animal derived in energy percent
```{r}
symnum.args <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))

#calculating animal and plant percent, per day

plant.animal <- merge(plant, animal, by = c("sample_day","source","Energie_bls"),all.x = TRUE, all.y = TRUE)
plant.animal <- subset(plant.animal,select = c('sample_day','source','Energie_bls'))

#sum of energy intake in one day

plant.animal$energy_sum_source <- with(plant.animal, ave(Energie_bls, sample_day, source, FUN = sum))

plant.animal$energy_sum_total <- with(plant.animal, ave(Energie_bls, sample_day, FUN = sum))

plant.animal$percent <- (100/plant.animal$energy_sum_total)*plant.animal$energy_sum_source

plant.animal$percent.plant <- with(plant.animal, ifelse(source=='plant', percent,0))
plant.animal$percent.animal <- with(plant.animal, ifelse(source=='animal', percent,0))

pa <- plant.animal %>% group_by(sample_day,percent.animal,percent.plant)

pa <- plant.animal  %>%
    group_by(sample_day) %>%
    summarise(percent.plant = unique(percent.plant[percent.plant!=0]), 
              percent.animal = unique(percent.animal[percent.animal!=0]))

pa$correct_percent <- pa$percent.plant+pa$percent.animal == 100

#coloring by U1 U2

pa$U1 <- grepl('U1',pa$sample_day)
pa$color <- ifelse(pa$U1 == 'TRUE','U1','U2')

#density plot

p <- ggplot(pa, aes(x=`percent.plant`, y=`percent.animal`, color=`color`)) +
  geom_point() + ggtitle('density plot PRE and POST nutritional protocols\nper day')
# + geom_smooth(method=lm) 

# ggplotly(p, tooltip = c('x','y','text'))
```


#abundance of plant and animal derived foods,per week
```{r}
#subset of k.bls_prodi per week

plant.animal.w <- merge(plant, animal, by = c("file_name","Phenotype","source","Energie_bls"),all.x = TRUE, all.y = TRUE)
plant.animal.w <- subset(plant.animal.w,select = c('file_name','Phenotype','source','Energie_bls'))

#sum of energy intake in one week

plant.animal.w$energy_sum_source <- with(plant.animal.w, ave(Energie_bls,file_name, source, FUN = sum))

plant.animal.w$energy_sum_total <- with(plant.animal.w, ave(Energie_bls, file_name, FUN = sum))


plant.animal.w$percent <- (100/plant.animal.w$energy_sum_total)*plant.animal.w$energy_sum_source

plant.animal.w$percent.plant <- with(plant.animal.w, ifelse(source=='plant', percent,0))
plant.animal.w$percent.animal <- with(plant.animal.w, ifelse(source=='animal', percent,0))

pa.w <- plant.animal.w %>% group_by(file_name,percent.animal,percent.plant)

pa.w <- plant.animal.w  %>%
    group_by(file_name) %>%
    summarise(percent.plant = unique(percent.plant[percent.plant!=0]), 
              percent.animal = unique(percent.animal[percent.animal!=0]))

pa.w$correct_percent <- pa.w$percent.plant+pa.w$percent.animal == 100

#coloring by U1 U2

pa.w$U1 <- grepl('U1',pa.w$file_name)
pa.w$color <- ifelse(pa.w$U1 == 'TRUE','U1','U2')

#density plot

p.w <- ggplot(pa.w, aes(x=`percent.plant`, y=`percent.animal`, color=`color`)) +
  geom_point() + ggtitle('density plot PRE and POST nutritional protocols')
# + geom_smooth(method=lm)

mean(pa.w$percent.plant)
sd(pa.w$percent.plant)

mean(pa.w$percent.animal)
sd(pa.w$percent.animal)

# ggplotly(p, tooltip = c('x','y','text'))
```

#stacked bar plot 
per week
```{r}
#percentage stacked bar plots

pa.w$count <- pa.w$percent.animal+pa.w$percent.plant
pa.w <- subset(pa.w,select = c(`file_name`,`percent.plant`,`percent.animal`,`count`))

#Melt and plot
perc.w <- pa.w %>% dplyr::select(-count) %>%
  pivot_longer(cols = -file_name) %>%
  ggplot(aes(x=file_name,y=value,fill=name)) + 
  geom_bar(position="fill", stat="identity")+
  theme(axis.text = element_text(angle = 90))

##only U1

pa.w.u1 <- pa.w[grepl(pattern = 'U1',pa.w$file_name), ]

summary(pa.w.u1$percent.animal)
sd(pa.w.u1$percent.animal)

perc.w.u1 <- pa.w.u1 %>% dplyr::select(-count) %>%
  pivot_longer(cols = -file_name) %>%
  ggplot(aes(x=file_name,y=value,fill=name)) + 
  geom_bar(position="fill", stat="identity")+
  theme(axis.text = element_text(angle = 90)) +ggtitle('animal and plant derived food\nPRE nutritional protocols')

##only U2

pa.w.u2 <- pa.w[grepl(pattern = 'U2',pa.w$file_name), ]

summary(pa.w.u2$percent.animal)
sd(pa.w.u2$percent.animal)

perc.w.u2 <- pa.w.u2 %>% dplyr::select(-count) %>%
  pivot_longer(cols = -file_name) %>%
  ggplot(aes(x=file_name,y=value,fill=name)) + 
  geom_bar(position="fill", stat="identity")+
  theme(axis.text = element_text(angle = 90)) +ggtitle('animal and plant derived food\nketo nutritional protocols')


#absolute stacked bar plots,per week

plant.animal.w.a <- merge(plant, animal, by = c("file_name","source"),all.x = TRUE, all.y = TRUE)
plant.animal.w.a <- subset(plant.animal.w,select = c('file_name','source'))

#count occurrences of plant and animal in plant.animal.w.a$source, conditioned by file_name

plant.animal.w.a <- plant.animal.w.a %>% group_by(file_name,source) %>% summarise(N=n()) %>% as.data.frame(.)
colnames(plant.animal.w.a) <- c('file_name','source','count')

##only U1

pa.w.a.u1 <- plant.animal.w.a[grepl(pattern = 'U1',plant.animal.w.a$file_name), ]

abs.U1 <- ggplot(pa.w.a.u1, aes(x = `file_name`,y = `count`))  + geom_bar(
    aes(fill = `source`),
    stat = "identity", position = position_stack())+ ggtitle('PRE protocols absolut number plant and animal foods')+ theme(axis.text = element_text(angle = 90))

##only U2

pa.w.a.u2 <- plant.animal.w.a[grepl(pattern = 'U2',plant.animal.w.a$file_name), ]

abs.U2 <- ggplot(pa.w.a.u2, aes(x = `file_name`,y = `count`))  + geom_bar(
    aes(fill = `source`),
    stat = "identity", position = position_stack())+ ggtitle('keto protocols absolut number plant and animal foods')+ theme(axis.text = element_text(angle = 90))


```

#boxplots,summarised
```{r}
#energy percent

pa.w$group <- ifelse(grepl('U2', pa.w$file_name), 'keto', 'non keto')


pp <- ggplot(pa.w,aes(x=group, y=percent.plant)) +
  geom_boxplot()+ylim(0,100)+ stat_compare_means(comparison = list(c('keto','non keto')),hide.ns = FALSE,symnum.args = symnum.args)+geom_jitter(shape = 21)

ap <- ggplot(pa.w,aes(x=group, y=percent.animal)) +
  geom_boxplot()+ylim(0,100)+ stat_compare_means(comparison = list(c('keto','non keto')),hide.ns = FALSE,symnum.args = symnum.args)+geom_jitter(shape = 21)

# pa.w.plant <- dplyr::select(pa.w,starts_with(c('percent.plant','group')))
# t.pa.w <- t.test(pa.w.plant)

# pdf(paste0(output_dir,'groups.animal.plant.pdf'))
groups.animal.plant <- ggarrange(pp,ap, 
          # labels = c("A", "B"),
          ncol = 2, nrow = 1)

# dev.off() 
```

#phenotype
are low and high converters eating different amounts of animal and plant derived foods?
```{r}
pa.pheno <- plant.animal.w %>% group_by(file_name,percent.animal,percent.plant)
pa.pheno <- subset(pa.pheno,select = c(`Phenotype`,`file_name`,`percent.plant`,`percent.animal`))

pa.pheno <- pa.pheno  %>%
  group_by(file_name,Phenotype) %>%
    summarise(percent.plant = unique(percent.plant[percent.plant!=0]), 
              percent.animal = unique(percent.animal[percent.animal!=0]))

pa.pheno$group <- ifelse(grepl('U2', pa.pheno$file_name), 'U2', 'U1')
pa.pheno$group <- ifelse(grepl('U2',pa.pheno$group),'keto','non-keto')

#remove not classified and NA

pa.pheno <- pa.pheno[pa.pheno$Phenotype != 'not classified', ] 
pa.pheno <- pa.pheno[complete.cases(pa.pheno), ]

##summary statistics
pa.pheno.low <- pa.pheno[grepl('low',pa.pheno$Phenotype),]
summary(pa.pheno.low$percent.animal)
sd(pa.pheno.low$percent.animal)

pa.pheno.high <- pa.pheno[grepl('high',pa.pheno$Phenotype),]
summary(pa.pheno.high$percent.animal)
sd(pa.pheno.high$percent.animal)

# pa.pheno.not <- pa.pheno[grepl('not',pa.pheno$Phenotype),]
# summary(pa.pheno.not$percent.animal)
# sd(pa.pheno.not$percent.animal)
# write_tsv(pa.pheno,paste0(table_dir,'plant.animal.processed.tsv'))

#boxplots

pheno.p <- ggplot(pa.pheno,aes(x=Phenotype, y=percent.plant, fill=group)) +
  geom_boxplot() +ylim(0,100)  + stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE)#
# + geom_signif(comparison = list(c('high converter','low converter')),map_signif_level = TRUE,test = 'wilcox.test')

pheno.a <- ggplot(pa.pheno,aes(x=Phenotype, y=percent.animal, fill=group)) +
  geom_boxplot()+ylim(0,100) + stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE)#
# + geom_signif(comparison = list(c('high converter','low converter')),map_signif_level = TRUE,test = 'wilcox.test')

pheno.ag <- ggplot(pa.pheno,aes(x=group, y=percent.animal, fill=Phenotype)) +
  geom_boxplot()+ylim(0,100) + geom_jitter(shape = 21)  + stat_compare_means(comparison = list(c('U1','U2')),hide.ns = FALSE)#
# + geom_signif(comparison = list(c('U1','U2')),map_signif_level = TRUE,test = 'wilcox.test')

pheno.pg <- ggplot(pa.pheno,aes(x=group, y=percent.plant, fill=Phenotype)) +
  geom_boxplot()+ylim(0,100)+geom_jitter(shape = 21) + stat_compare_means(comparison = list(c('U1','U2')),hide.ns = FALSE)#p.format
# + geom_signif(comparison = list(c('U1','U2')),map_signif_level = TRUE,test = 'wilcox.test')

#only U1
 symnum.args <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))
 
pa.pheno.u1 <- subset(pa.pheno,grepl("U1",file_name))

pheno.u1.p <- ggplot(pa.pheno.u1,aes(x=Phenotype, y=percent.plant)) +
  geom_boxplot()+ylim(0,100) + stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE,symnum.args = symnum.args)+ggtitle('non-keto diet')
#
# +geom_signif(comparison = list(c('high converter','low converter')),map_signif_level = TRUE,test = 'wilcox.test')
  

pheno.u1.a <- ggplot(pa.pheno.u1,aes(x=Phenotype, y=percent.animal)) +
  geom_boxplot()+ylim(0,100) + stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE,symnum.args = symnum.args)+ggtitle('non-keto diet')
#
# +geom_signif(comparison = list(c('high converter','low converter')),map_signif_level = TRUE,test = 'wilcox.test')

#only U2

pa.pheno.u2 <- subset(pa.pheno,grepl("U2",file_name))

pheno.u2.p <- ggplot(pa.pheno.u2,aes(x=Phenotype, y=percent.plant)) +
  geom_boxplot()+ylim(0,100)+ stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE,symnum.args = symnum.args)+ggtitle('keto diet')
#
  # +geom_signif(comparison = list(c('high converter','low converter')),map_signif_level = TRUE, test = 'wilcox.test')


pheno.u2.a <- ggplot(pa.pheno.u2,aes(x=Phenotype, y=percent.animal)) +
  geom_boxplot()+ylim(0,100) + stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE,symnum.args = symnum.args)+ggtitle('keto diet')
#
# +geom_signif(comparison = list(c('high converter','low converter')),map_signif_level = TRUE, test = 'wilcox.test')

phenotype.group.animal.plant <- ggarrange(pheno.u1.p,pheno.u1.a,pheno.u2.p,pheno.u2.a, 
          labels = c("A", "B","C","D"),
          ncol = 2, nrow = 2)

#entire data high and low converters

pheno.aa <- ggplot(pa.pheno,aes(x=Phenotype, y=percent.animal)) +
  geom_boxplot()+ylim(0,100)+geom_jitter(shape = 21)+ stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE,symnum.args = symnum.args)

# + stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = TRUE)#

pheno.ap <- ggplot(pa.pheno,aes(x=Phenotype, y=percent.plant)) +
  geom_boxplot()+ylim(0,100)+geom_jitter(shape = 21)+ stat_compare_means(comparison = list(c('high converter','low converter')),hide.ns = FALSE,symnum.args = symnum.args) #

phenotype.animal.plant <- ggarrange(pheno.aa,pheno.ap,ap,pp, 
          labels = c("A","B","C","D"),
          ncol = 2, nrow = 2)

```

###export
```{r}
# pdf(paste0(output_dir,'diet.protocols/animal.plant.pdf'))
# pdf(paste0(diet.protocols,'animal.plant.pdf'))
groups.animal.plant 

phenotype.animal.plant

phenotype.group.animal.plant

# dev.off()

# write_tsv(plant.animal.w,paste0(table_dir,'plant.animal.w.tsv'))
# write_tsv(plant.animal.w,paste0(table_dir,'plant.animal.percent.tsv'))
```

