---
title: "UniFrac"
output: html_notebook
---


#libraries
```{r}
library(tidyverse)
library(ape)
library(phyloseq)
library(ggplot2)
library(devtools)
library(factoextra)
# library(GMDecomp)
library(apcluster)
library(plotly)
# library(RFLPtools)
# devtools::install_github("taryue/GMDecomp")

library(Rtsne)
library(FactoMineR)
library(rgr)

library(ggdendro)

# library(DECIPHER)

library(ggpubr)
```

# Paths
```{r}
wd <- '../../' 

output_dir = paste0(wd,'results/')
diet.protocols = paste0(output_dir,'diet.protocols/')

data_dir = paste0(wd,'data/')
table_dir = paste0(data_dir, 'tables/')
biom_dir = paste0(data_dir,'silva.feces/')
```

## Create Output Dirs
```{r}
dir.create(output_dir)
dir.create(table_dir)
```

# Data Import
```{r}
keto.bls <- read_tsv(paste0(table_dir, 'keto.bls_04.08.tsv'))
bls.extended <- read_tsv(paste0(table_dir,'bls.extended_01.08.tsv'))
k.bls_prodi <- read_tsv(paste0(table_dir,'k.bls_prodi_13_08.tsv'))

pca.bls <- read_tsv(paste0(table_dir,'pca.bls_14.10.tsv'))
nutr.sum <- read_tsv(paste0(table_dir,'nutr.sum_14.10.tsv'))

# #trees
tree.conv <- read.tree(paste0(table_dir,'tree.conv.19.08.tre'))
tree.keto.bls <- read.tree(paste0(table_dir,'tree.keto.bls.19.08.tre'))

```



generating phylogenetic object
keto samples, conventional clusters

manually assign distances according to 
groups and subgroups that are already defined by the bls

```{r fig.height=10}
set.seed(73)
keto.bls <- keto.bls[!(keto.bls$item_cat == 'Gewürze, Würzmittel, Hilfsstoffe' ),]##| keto.bls$item_cat =='Menükomponenten'

tmp <- unique(subset(keto.bls,select = c(`item_cat`,`item_subcat`,`item_prodi`))) %>% distinct(.)
r <- tmp
colnames(r) <- c('item_cat.select','item_subcat.select','item_prodi.select') 


r <- tidyr::uncount(r, dplyr::n())
# tmp <- tmp[rep(seq_len(nrow(tmp)), nrow(r)), ]
tmp <- purrr::map_dfr(seq_len(1250), ~tmp)##nrow(tmp)/nrow(r)=1044  ##w/o gewürze 1280
tmp <- cbind(tmp,r)

cat.dist <-
  tmp %>% 
  mutate(same_subcat = item_subcat.select == item_subcat,
         same_cat = item_cat.select == item_cat) %>% 
  mutate(total = (as.numeric(!same_cat) + as.numeric(!same_subcat)) / 2) %>% 
  dplyr::select(item_prodi.select, item_prodi, total) %>%
  
  pivot_wider(names_from = 'item_prodi.select', values_from = 'total') %>% 
  column_to_rownames('item_prodi') %>%
  as.dist(.)

cat.dist[is.na(cat.dist)] <- 0

#cats should all have own node
#respective sub_cats too
```

#phylo object conv. cluster
```{r fig.height=10}
d.cat_data <- hclust(cat.dist, method = 'complete')
dend.cat <- as.dendrogram(d.cat_data)
tree.conv <- ape::as.phylo(d.cat_data,use.labels = keto.bls$item_cat)


plot(tree.conv, type= 'fan')

# write.tree(tree.conv, file=paste0(table_dir,'tree.conv.19.08.tre',append = FALSE))


```
for UniFrac you need a phylogenetic object (--> nutrient clustering derived from notebook BLS.betadiv via PCA, convetional clustering see above)

matrix for UniFrac weighted and unweighted food-samples df

18.08
#conventional clusters UniFrac
one sample one week
```{r}
k.bls_prodi$file_name <- gsub(k.bls_prodi$file_name,pattern='u',replacement = 'U') %>% gsub(pattern = '_',replacement = '.')
k.bls_prodi <- k.bls_prodi[k.bls_prodi$database == 'bls',]
k.bls_prodi$Sex[k.bls_prodi$Sex == FALSE] <- 'female'
k.bls_prodi$Sex[is.na(k.bls_prodi$Sex)] <- 0
k.bls_prodi$Sex[k.bls_prodi$Sex == 0] <- 'male'


##weighted UniFrac

df.unifrac <- subset(k.bls_prodi,select=c(`file_name`,`item`,`Energie_bls`,`Energie (kcal)`))


df.unifrac$calculation <- (100/df.unifrac$`Energie_bls`)*df.unifrac$`Energie (kcal)`

##calculation for weighted UniFrac

df.unifrac <- as.data.frame(df.unifrac)
df.unifrac$calculation[is.nan(df.unifrac$calculation)] <-  0

#otu_table generation for weighted UniFrac
otu.unifrac <- subset(df.unifrac,select = c(`item`,`calculation`,`file_name`))

otu.unifrac <- otu.unifrac  %>% group_by(file_name,item) %>% summarize(sum = sum(calculation)) %>% pivot_wider(names_from = 'item',values_from = 'sum') %>% as.data.frame(.)

otu.unifrac[is.na(otu.unifrac)] <- 0

# write_tsv(otu.unifrac,path=paste0(table_dir,'otu.conv.w.19.08.tsv',append=FALSE))

row.names(otu.unifrac) <- otu.unifrac$file_name 
otu.unifrac <- otu.unifrac[ ,-1]


otu.weighted <- otu_table(otu.unifrac,taxa_are_rows = FALSE)

#weighted UniFrac
phylo.conv.w <- phyloseq(otu.weighted,tree.conv)
unifrac.keto.conv.w <- UniFrac(phylo.conv.w,weighted = TRUE,fast = TRUE)
pcoa_keto.w <- ape::pcoa(unifrac.keto.conv.w)

##unweighted UniFrac

sub.bell <- k.bls_prodi[!(k.bls_prodi$item_descr=="gemüsepaprika rot"),]
sub.nudels <- k.bls_prodi[!(k.bls_prodi$item_descr=="nudeln (4)"),]

unifrac <- subset(k.bls_prodi,select=c(`file_name`,`item`))

unifrac <- as.data.frame(unifrac)
unifrac <- xtabs(~ item + file_name , data = unifrac)##how often consumed/d

unifrac <- as.data.frame(unifrac)  
unifrac <- pivot_wider(data = unifrac, names_from = 'item', values_from = 'Freq') %>% as.data.frame(.)

# write_tsv(unifrac,path=paste0(table_dir,'otu.conv.uw.19.08.tsv',append=FALSE,col_names = TRUE))

row.names(unifrac) <- unifrac$file_name
unifrac <- unifrac[ ,-1]

unifrac <- otu_table(unifrac,taxa_are_rows = FALSE)


#unweighted UniFrac
phylo.conv.uw <- phyloseq(unifrac,tree.conv)
unifrac.keto.conv.uw <- UniFrac(phylo.conv.uw,weighted = FALSE,fast = TRUE)
pcoa_keto.uw <- ape::pcoa(unifrac.keto.conv.uw)

get_explained_variance <- function(pcoa){
  return(paste0(round(pcoa$values$Rel_corr_eig *100, 2), '%'))
}

get_explained_variance(pcoa_keto.uw)[1:2]##"11.21%" "8.45%"
get_explained_variance(pcoa_keto.w)[1:2]##"12.69%" "9.66%" 

#varinace
pcoa_keto.uw$vectors <- as.data.frame(pcoa_keto.uw$vectors)
pcoa_keto.w$vectors <- as.data.frame(pcoa_keto.w$vectors)

varconvAxis.1 <- var(pcoa_keto.uw$vectors$Axis.1)##0.02851293
varconvAxis.2 <- var(pcoa_keto.uw$vectors$Axis.2)##0.01470678

varconvwAxis.1 <- var(pcoa_keto.w$vectors$Axis.1)##0.04093632
varconvwAxis.2 <- var(pcoa_keto.w$vectors$Axis.2)##0.03483469

##visualization

#Phenotype
p.metadata.conv.uw <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
conv.uw.pheno <- pcoa_keto.uw$vectors %>%  #or pcoa_keto.w/pcoa_nutr.w or pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., p.metadata.conv.uw, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA unweighted UniFrac conventional clusters')+ theme(plot.title = element_text(size=10))

#Sex
s.metadata.conv.uw <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Sex) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
conv.uw.sex <- pcoa_keto.uw$vectors %>%  #or pcoa_keto.w/pcoa_nutr.w or pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., s.metadata.conv.uw, by = 'Name') %>%
  mutate(sex = ifelse(Sex == 'not classified', NA, Sex)) %>% 
  # mutate(sex = ifelse(Sex == 'female', FALSE, Sex)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Sex),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA unweighted UniFrac conventional clusters')+ theme(plot.title = element_text(size=10))##FALSE female


#Phenotype
p.metadata.conv.w <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
conv.w.pheno <- pcoa_keto.w$vectors %>%  #/pcoa_nutr.w or pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., p.metadata.conv.w, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA weighted UniFrac conventional clusters')+ theme(plot.title = element_text(size=10))

#Sex
s.metadata.conv.w <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Sex) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
conv.w.sex <- pcoa_keto.w$vectors %>%  #/pcoa_nutr.w or pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., s.metadata.conv.w, by = 'Name') %>%
  mutate(Pheno = ifelse(Sex == 'not classified', NA, Sex)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Sex),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA weighted UniFrac conventional clusters')+ theme(plot.title = element_text(size=10))



```
#conventional cluster
one sample one day
```{r}
#unweighted UniFrac per day
otu.unweighted.samp <- subset(k.bls_prodi,select=c(`file_name`,`item`,`day`))

otu.unweighted.samp$sample_day <- paste(otu.unweighted.samp$file_name,otu.unweighted.samp$day)
otu.unweighted.samp <- otu.unweighted.samp[ ,-c(1)]
otu.unweighted.samp <- as.data.frame(otu.unweighted.samp)
otu.unweighted.samp <- xtabs(~ item + sample_day , data = otu.unweighted.samp)##how often consumed/d

otu.unweighted.samp <- as.data.frame(otu.unweighted.samp)  
otu.unweighted.samp <- pivot_wider(data = otu.unweighted.samp, names_from = 'item', values_from = 'Freq') %>% as.data.frame(.)

# write_tsv(otu.unweighted.samp,path=paste0(table_dir,'otu.conv.uw.samp.19.08.tsv',append=FALSE))

row.names(otu.unweighted.samp) <- otu.unweighted.samp$sample_day
otu.unweighted.samp <- otu.unweighted.samp[ ,-1]
otu.unweighted.samp <- otu_table(otu.unweighted.samp,taxa_are_rows = FALSE)

phylo.conv.uw.samp <- phyloseq(otu.unweighted.samp,tree.conv)
unifrac.keto.conv.uw.samp <- UniFrac(phylo.conv.uw.samp,weighted = FALSE,fast = TRUE)
pcoa_keto.uw.samp <- ape::pcoa(unifrac.keto.conv.uw.samp)

##visualization

#Phenotype
p.metadata.conv.uw.samp <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  rename(Name = file_name)
pcoa_keto.uw.samp$vectors %>%  #or pcoa_keto.w/pcoa_nutr.w or pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., p.metadata.conv.uw.samp, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA unweighted UniFrac conventional clusters')


#weighted UniFrac per day

df.unifrac.samp <- subset(k.bls_prodi,select=c(`file_name`,`item`,`Energie_bls`,`Energie (kcal)`,`day`))

##calculation for weighted UniFrac
df.unifrac.samp$calculation <- apply(X=df.unifrac.samp,MARGIN=2,FUN=function(line){(100/df.unifrac$`Energie_bls`)*df.unifrac$`Energie (kcal)`})
df.unifrac.samp <- as.data.frame(df.unifrac.samp)
df.unifrac.samp$sample_day <- paste(df.unifrac.samp$file_name,df.unifrac.samp$day)
df.unifrac.samp <- df.unifrac.samp[ ,-c(1)]

#otu_table generation for weighted UniFrac
otu.unifrac.samp <- subset(df.unifrac.samp,select = c(`item`,`calculation`,`sample_day`))

otu.unifrac.samp <- otu.unifrac.samp %>% group_by(item,sample_day) %>% summarize(sum = sum(calculation)) %>% pivot_wider(names_from = 'item',values_from = 'sum') %>% as.data.frame(.)

otu.unifrac.samp[is.na(otu.unifrac.samp)] <- 0
row.names(otu.unifrac.samp) <- otu.unifrac.samp$sample_day
otu.unifrac.samp <- otu.unifrac.samp[ ,-1]


otu.weighted.samp <- otu_table(otu.unifrac.samp,taxa_are_rows = FALSE)

phylo.conv.w.samp <- phyloseq(otu.weighted.samp,tree.conv)
unifrac.keto.conv.w.samp <- UniFrac(phylo.conv.w.samp,weighted = FALSE,fast = TRUE)
pcoa_keto.w.samp <- ape::pcoa(unifrac.keto.conv.w.samp)

##visualization

#Phenotype
p.metadata.conv.w.samp <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  rename(Name = file_name)
pcoa_keto.w.samp$vectors %>%  ##error
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., p.metadata.conv.w.samp, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA weighted UniFrac conventional clusters')

```

19.08
#nutrient based cluster UniFrac
one sample one week
```{r}
##tree.keto.bls 

##weighted UniFrac
phylo.nutr.w <- phyloseq(otu.weighted,tree.keto.bls)
unifrac.nutr.w <- UniFrac(phylo.nutr.w,weighted = TRUE,fast = TRUE)
pcoa_nutr.w <- ape::pcoa(unifrac.nutr.w)

##unweighted UniFrac
phylo.nutr.uw <- phyloseq(unifrac,tree.keto.bls)
unifrac.nutr.uw <- UniFrac(phylo.nutr.uw,weighted = FALSE,fast = TRUE)
pcoa_nutr.uw <- ape::pcoa(unifrac.nutr.uw)

#visualizations

#Phenotype
metadata.nutr.w <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
nutr.w.pheno <- pcoa_nutr.w$vectors %>%  #/pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., metadata.nutr.w, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA weighted UniFrac nutrient clusters')+ theme(plot.title = element_text(size=10))


#Sex
metadata.nutr.w <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Sex) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
nutr.w.sex <- pcoa_nutr.w$vectors %>%  #/pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., metadata.nutr.w, by = 'Name') %>%
  mutate(Pheno = ifelse(Sex == 'not classified', NA, Sex)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Sex),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA weighted UniFrac nutrient clusters')+ theme(plot.title = element_text(size=10))

#Phenotype
metadata.nutr.uw <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
nutr.uw.pheno <- pcoa_nutr.uw$vectors %>%  #
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., metadata.nutr.uw, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA unweighted UniFrac nutrient clusters')+ theme(plot.title = element_text(size=10))

# ggplotly(ug,tooltip = c('x','y','text'))

#Sex
metadata.nutr.uw <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Sex) %>% 
  distinct() %>% 
  dplyr::rename(Name = file_name)
nutr.uw.sex <- pcoa_nutr.uw$vectors %>%  #/pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., metadata.nutr.uw, by = 'Name') %>%
  mutate(Pheno = ifelse(Sex == 'not classified', NA, Sex)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Sex),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA unweighted UniFrac nutrient clusters')+ theme(plot.title = element_text(size=10))

#variance

get_explained_variance(pcoa_nutr.uw)[1:2]
get_explained_variance(pcoa_nutr.w)[1:2]

pcoa_nutr.uw$values$Relative_eig

pcoa_nutr.uw$vectors <- as.data.frame(pcoa_nutr.uw$vectors)
pcoa_nutr.w$vectors <- as.data.frame(pcoa_nutr.w$vectors)

varAxis.1 <- var(pcoa_nutr.uw$vectors$Axis.1)##0.005039151
# -->100*0.02635798=2.635798% WRONG?
varAxis.2 <- var(pcoa_nutr.uw$vectors$Axis.2)##0.003714742
#--> 100*0.02117861=2.117861% WRONG?
#very little variance on both axis

varwAxis.1 <- var(pcoa_nutr.w$vectors$Axis.1)##0.04277659
varwAxis.2 <- var(pcoa_nutr.w$vectors$Axis.2)##0.0120085

var_pop <- function(x) {       # Create function for population variance
  mean((x - mean(x))^2)
}

var_popAxis.1 <- var_pop(pcoa_nutr.uw$vectors$Axis.1)##0.02603901
#--> 100*0.02603901=2.603901% WRONG?
var_popAxis.2 <- var_pop(pcoa_nutr.uw$vectors$Axis.2)##0.02078877
#--> 100*0.02078877=2.078877% WRONG? 

# ggplotly(k,tooltip = c('x','y','text'))

```

#nutrient cluster
one sample one day
```{r}
#unweighted UniFrac per day

phylo.nutr.uw.samp <- phyloseq(otu.unweighted.samp,tree.keto.bls)
unifrac.nutr.uw.samp <- UniFrac(phylo.nutr.uw.samp,weighted = FALSE,fast = TRUE)
pcoa_nutr.uw.samp <- ape::pcoa(unifrac.nutr.uw.samp)

##visualization

#Phenotype
p.metadata.nutr.uw.samp <- 
  k.bls_prodi %>% 
  dplyr::select(file_name:Phenotype) %>% 
  distinct() %>% 
  rename(Name = file_name)
pcoa_nutr.uw.samp$vectors %>%  #or pcoa_keto.w/pcoa_nutr.w or pcoa_nutr.uw
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  dplyr::select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., p.metadata.nutr.uw.samp, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA unweighted UniFrac nutrient clusters')

#weighted UniFrac
phylo.nutr.w.samp <- phyloseq(otu.weighted.samp,tree.keto.bls)
unifrac.nutr.w.samp <- UniFrac(phylo.nutr.w.samp,weighted = FALSE,fast = TRUE)
pcoa_nutr.w.samp <- ape::pcoa(unifrac.nutr.w.samp)

##visualization

#Phenotype
p.metadata.nutr.w.samp <- 
  k.bls_prodi %>% 
  select(file_name:Phenotype) %>% 
  distinct() %>% 
  rename(Name = file_name)
pcoa_nutr.w.samp$vectors %>%  ##error
  as.data.frame(.) %>% 
  rownames_to_column('Name') %>% 
  select(Name, Axis.1:Axis.5) %>%
  separate(Name, remove = F, into = c('Sample','U'), sep = '[.]') %>% 
  left_join(., p.metadata.nutr.w.samp, by = 'Name') %>%
  mutate(Pheno = ifelse(Phenotype == 'not classified', NA, Phenotype)) %>% 
  ggplot(aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill = U, alpha = Pheno),
    shape = 21, col = 'black', size = 4) + ggtitle('PcoA weighted UniFrac nutrient clusters')
```

###export
```{r}
pdf(paste0(diet_dir,'unifrac.pc.pdf'))

ggarrange(conv.uw.pheno,conv.uw.sex, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,heights = c(1,1))

ggarrange(nutr.uw.pheno,nutr.uw.sex, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,heights = c(1,1))

ggarrange(conv.w.pheno,conv.w.sex, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,heights = c(1,1))

ggarrange(nutr.w.pheno,nutr.w.sex, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,heights = c(1,1))

dev.off()
```



