---
title: "correlation rarefied taxa with nutrients (all,U1,U2)"
output: html_notebook
---

#libraries
```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("biomformat")

# if (!requireNamespace('BiocManager', quietly = TRUE))
#     install.packages('BiocManager')
# 
#  BiocManager::install('EnhancedVolcano')

# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("a4Base")
  

library(tidyverse)
library(biomformat)
# library(data.table)
library(compositions)
# library(corrplot)
# library(EnhancedVolcano)
library(a4Base)
library(plotly)
library(ggpubr)

library(Maaslin2)
```

# Paths
```{r}
wd <- '../../' 

output_dir = paste0(wd,'results/')
diet.protocols = paste0(output_dir,'diet.protocols/')

data_dir = paste0(wd,'data/')
table_dir = paste0(data_dir, 'tables/')
biom_dir = paste0(data_dir,'silva.feces/')
```

## Create Output Dirs
```{r}
dir.create(output_dir)
dir.create(table_dir)
```

# Data Import
```{r}

biom.L6 <- read.csv(paste0(biom_dir, 'rarefied1623_filtered300_L6.csv'),header = TRUE,sep = ';',row.names = 1)
biom.L7 <- read.csv(paste0(biom_dir, 'rarefied1623_filtered300_L7.csv'),header = TRUE,sep = ';',row.names = 1)

nutr.sum.z <- read_tsv(paste0(table_dir, 'norm.keto.bls.17.02.tsv'))%>% column_to_rownames(.,'file_name')

```

##biom
pseudo count
relative abundance
centered log ratio
```{r}
##preparation

#two samples for U1,U2 and U3
#merging them ore deleting?

#L6 genus
biom.L6 <- biom.L6[!grepl('.2U',row.names(biom.L6)),]
biom.L6 <- biom.L6[!grepl(pattern = 'U3',row.names(biom.L6)),]
biom.L6 <- dplyr::select(biom.L6,everything(),-starts_with(c('Sex','Time','Proband','Body','Day')))

row.names(biom.L6) <- str_replace(row.names(biom.L6),pattern = 'ST.',replacement = '') %>% 
  str_replace(.,pattern = '0U', replacement = 'U')

#species L7
biom.L7 <- biom.L7[!grepl('.2U',row.names(biom.L7)),]
biom.L7 <- biom.L7[!grepl(pattern = 'U3',row.names(biom.L7)),]
biom.L7 <- dplyr::select(biom.L7,everything(),-starts_with(c('Sex','Time','Proband','Body','Day')))

row.names(biom.L7) <- str_replace(row.names(biom.L7),pattern = 'ST.',replacement = '') %>% 
  str_replace(.,pattern = '0U', replacement = 'U')

#L6  relative abundance, clr

biom.L6.sub <- data.matrix(biom.L6)
biom.L6.sub <- biom.L6.sub+0.5
rowsum <- rowSums(biom.L6.sub)

biom.L6 <- cbind(biom.L6.sub,rowsum)

#OTU count / OTU sample sum = rel. abund  taxa

biom.L6 <- as.data.frame(biom.L6)
biom.L6.rel <- biom.L6/biom.L6$rowsum
biom.L6.rel <- dplyr::select(biom.L6.rel,everything(),-starts_with('row'))

#L7  relative abundance, clr

biom.L7.sub <- data.matrix(biom.L7)
biom.L7.sub <- biom.L7.sub+0.5
rowsum <- rowSums(biom.L7.sub)

biom.L7 <- cbind(biom.L7.sub,rowsum)

#OTU count / OTU sample sum = rel. abund  taxa

biom.L7 <- as.data.frame(biom.L7)
biom.L7.rel <- biom.L7/biom.L7$rowsum
biom.L7.rel <- dplyr::select(biom.L7.rel,everything(),-starts_with('row'))

# biom.L6.rel <- rownames_to_column(biom.L6.rel)
# write_tsv(biom.L6.rel,paste0(table_dir,'rarefied.rel.L6.tsv'))

# biom.L7.rel <- rownames_to_column(biom.L7.rel)
# write_tsv(biom.L7.rel,paste0(table_dir,'rarefied.rel.L7.tsv'))
```

clr transformation
```{r}
biom.L6.rel.clr <- clr(biom.L6.rel) %>% as.data.frame(.)

biom.L7.rel.clr <- clr(biom.L7.rel) %>% as.data.frame(.)

rownames_to_column(biom.L6.rel.clr) %>%
write_tsv(.,paste0(table_dir,'biom.L6.rel.clr_19.02.tsv'))

rownames_to_column(biom.L7.rel.clr) %>%
write_tsv(.,paste0(table_dir,'biom.L7.rel.clr_19.02.tsv'))
```

subsetting biom.rel.clr and nutr.sum according to day U1,U2
```{r}
##U1 
biom.L6.rel.clr.U1 <- biom.L6.rel.clr[grepl('U1',row.names(biom.L6.rel.clr)),]

biom.L7.rel.clr.U1 <- biom.L7.rel.clr[grepl('U1',row.names(biom.L7.rel.clr)),]

nutr.sum.z.U1 <- nutr.sum.z[grepl('U1',row.names(nutr.sum.z)),]

rownames_to_column(biom.L6.rel.clr.U1) %>% write_tsv(.,paste0(table_dir,'biom.L6.rel.clr.U1_19.02.tsv'))

rownames_to_column(biom.L7.rel.clr.U1) %>%
write_tsv(.,paste0(table_dir,'biom.L7.rel.clr.U1_19.02.tsv'))

##U2 
biom.L6.rel.clr.U2 <- biom.L6.rel.clr[grepl('U2',row.names(biom.L6.rel.clr)),]

biom.L7.rel.clr.U2 <- biom.L7.rel.clr[grepl('U2',row.names(biom.L7.rel.clr)),]

nutr.sum.z.U2 <- nutr.sum.z[grepl('U2',row.names(nutr.sum.z)),]

rownames_to_column(biom.L6.rel.clr.U2) %>%
write_tsv(.,paste0(table_dir,'biom.L6.rel.clr.U2_19.02.tsv'))

rownames_to_column(biom.L7.rel.clr.U2) %>%
write_tsv(.,paste0(table_dir,'biom.L7.rel.clr.U2_19-02.tsv'))
```

###L6
##correlation between nutrients and biom.L6
merge taxa table and nutrient table (standardized) together
correlate every nutrient with every taxa
spearman correlation
-log10(BH adjusted p.values)

nutr.sum.z kommt von biom.volcano
```{r}

merge.biom.L6.nutr <- biom.L6.rel.clr %>% merge(.,nutr.sum.z,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L6.nutr[is.na(merge.biom.L6.nutr)] <- 0

merge.biom.L6.nutr$file_name <- row.names(merge.biom.L6.nutr)
write_tsv(merge.biom.L6.nutr,paste0(table_dir,'corr.biom.L6.19.02.tsv',append=FALSE))
merge.biom.L6.nutr <- select(merge.biom.L6.nutr,everything(),-starts_with('file_name'))

cor.test.biom.L6 <- merge.biom.L6.nutr %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Energie (kcal)`:`Energie, Gesamt (kcal)`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L6[is.na(cor.test.biom.L6)] <- 0


p.adj.biom.L6 <- subset(cor.test.biom.L6,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L6 <- subset(cor.test.biom.L6,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L6 <- log2(r.cor.biom.L6) %>% as.numeric(.)
cor.test.log.biom.L6[is.nan(cor.test.log.biom.L6)] <- 0
cor.test.log.biom.L6[is.infinite(cor.test.log.biom.L6)] <- 0
cor.test.log.biom.L6 <- as.data.frame(cor.test.log.biom.L6)

r.cor.biom.L6[is.nan(r.cor.biom.L6)] <- 0
r.cor.biom.L6[is.infinite(r.cor.biom.L6)] <- 0
r.cor.biom.L6 <- as.data.frame(r.cor.biom.L6)

#y axis -log10 p.value

p.adj.log.biom.L6 <- -log10(p.adj.biom.L6)
p.adj.log.biom.L6[is.infinite(p.adj.log.biom.L6)] <- 0
p.adj.log.biom.L6 <- as.data.frame(p.adj.log.biom.L6)

cor.test.adj.biom.L6 <- cbind(cor.test.biom.L6[ ,c(1,3)],p.adj.log.biom.L6) %>% cbind(.,r.cor.biom.L6) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L6) <- c('genus','nutrient','p.adj.log','estimate')

cor.test.adj.biom.L6 <- cor.test.adj.biom.L6[cor.test.adj.biom.L6$p.adj.log != 0 & cor.test.adj.biom.L6$estimate != 0,]

write_tsv(cor.test.adj.biom.L6,paste0(table_dir,'corr.nutr.biom.L6.tsv'))

```

###U1 correlation between nutrients and biom.L6.U1
```{r}

merge.biom.L6.nutr.U1 <- biom.L6.rel.clr.U1 %>% merge(.,nutr.sum.z.U1,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L6.nutr.U1[is.na(merge.biom.L6.nutr.U1)] <- 0

merge.biom.L6.nutr.U1$file_name <- row.names(merge.biom.L6.nutr.U1)
write_tsv(merge.biom.L6.nutr.U1,paste0(table_dir,'corr.biom.L6.U1.19.02.tsv',append=FALSE))
merge.biom.L6.nutr.U1 <- select(merge.biom.L6.nutr.U1,everything(),-starts_with('file_name'))

cor.test.biom.L6.U1 <- merge.biom.L6.nutr.U1 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Energie (kcal)`:`Energie, Gesamt (kcal)`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L6.U1[is.na(cor.test.biom.L6.U1)] <- 0


p.adj.biom.L6.U1 <- subset(cor.test.biom.L6.U1,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L6.U1 <- subset(cor.test.biom.L6.U1,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L6.U1 <- log2(r.cor.biom.L6.U1) %>% as.numeric(.)
cor.test.log.biom.L6.U1[is.nan(cor.test.log.biom.L6.U1)] <- 0
cor.test.log.biom.L6.U1[is.infinite(cor.test.log.biom.L6.U1)] <- 0
cor.test.log.biom.L6.U1 <- as.data.frame(cor.test.log.biom.L6.U1)

r.cor.biom.L6.U1[is.nan(r.cor.biom.L6.U1)] <- 0
r.cor.biom.L6.U1[is.infinite(r.cor.biom.L6.U1)] <- 0
r.cor.biom.L6.U1 <- as.data.frame(r.cor.biom.L6.U1)

#y axis -log10 p.value

p.adj.log.biom.L6.U1 <- -log10(p.adj.biom.L6.U1)
p.adj.log.biom.L6.U1[is.infinite(p.adj.log.biom.L6.U1)] <- 0
p.adj.log.biom.L6.U1 <- as.data.frame(p.adj.log.biom.L6.U1)

cor.test.adj.biom.L6.U1 <- cbind(cor.test.biom.L6.U1[ ,c(1,3)],p.adj.log.biom.L6.U1) %>% cbind(.,r.cor.biom.L6.U1) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L6.U1) <- c('genus','nutrient','p.adj.log','estimate')

cor.test.adj.biom.L6.U1 <- cor.test.adj.biom.L6.U1[cor.test.adj.biom.L6.U1$p.adj.log != 0 & cor.test.adj.biom.L6.U1$estimate != 0,]

write_tsv(cor.test.adj.biom.L6.U1,paste0(table_dir,'corr.nutr.biom.L6.U1.tsv'))

```

###U2 correlation between nutrients and biom.L6.U1
```{r}

merge.biom.L6.nutr.U2 <- biom.L6.rel.clr.U2 %>% merge(.,nutr.sum.z.U2,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L6.nutr.U2[is.na(merge.biom.L6.nutr.U2)] <- 0

merge.biom.L6.nutr.U2$file_name <- row.names(merge.biom.L6.nutr.U2)
write_tsv(merge.biom.L6.nutr.U2,paste0(table_dir,'corr.biom.L6.U2.19.02.tsv',append=FALSE))
merge.biom.L6.nutr.U2 <- select(merge.biom.L6.nutr.U2,everything(),-starts_with('file_name'))

cor.test.biom.L6.U2 <- merge.biom.L6.nutr.U2 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Energie (kcal)`:`Energie, Gesamt (kcal)`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L6.U2[is.na(cor.test.biom.L6.U2)] <- 0


p.adj.biom.L6.U2 <- subset(cor.test.biom.L6.U2,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L6.U2 <- subset(cor.test.biom.L6.U2,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L6.U2 <- log2(r.cor.biom.L6.U2) %>% as.numeric(.)
cor.test.log.biom.L6.U2[is.nan(cor.test.log.biom.L6.U2)] <- 0
cor.test.log.biom.L6.U2[is.infinite(cor.test.log.biom.L6.U2)] <- 0
cor.test.log.biom.L6.U2 <- as.data.frame(cor.test.log.biom.L6.U2)

r.cor.biom.L6.U2[is.nan(r.cor.biom.L6.U2)] <- 0
r.cor.biom.L6.U2[is.infinite(r.cor.biom.L6.U2)] <- 0
r.cor.biom.L6.U2 <- as.data.frame(r.cor.biom.L6.U2)

#y axis -log10 p.value

p.adj.log.biom.L6.U2 <- -log10(p.adj.biom.L6.U2)
p.adj.log.biom.L6.U2[is.infinite(p.adj.log.biom.L6.U2)] <- 0
p.adj.log.biom.L6.U2 <- as.data.frame(p.adj.log.biom.L6.U2)

cor.test.adj.biom.L6.U2 <- cbind(cor.test.biom.L6.U2[ ,c(1,3)],p.adj.log.biom.L6.U2) %>% cbind(.,r.cor.biom.L6.U2) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L6.U2) <- c('genus','nutrient','p.adj.log','estimate')

cor.test.adj.biom.L6.U2 <- cor.test.adj.biom.L6.U2[cor.test.adj.biom.L6.U2$p.adj.log != 0 & cor.test.adj.biom.L6.U2$estimate != 0,]

write_tsv(cor.test.adj.biom.L6.U2,paste0(table_dir,'corr.nutr.biom.L6.U2.tsv'))


```

###L7
##correlation between nutrients and biom.L7
```{r}

merge.biom.L7.nutr <- biom.L7.rel.clr %>% merge(.,nutr.sum.z,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L7.nutr[is.na(merge.biom.L7.nutr)] <- 0

merge.biom.L7.nutr$file_name <- row.names(merge.biom.L7.nutr)
write_tsv(merge.biom.L7.nutr,paste0(table_dir,'corr.biom.L7.19.02.tsv',append=FALSE))
merge.biom.L7.nutr <- select(merge.biom.L7.nutr,everything(),-starts_with('file_name'))

cor.test.biom.L7 <- merge.biom.L7.nutr %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter.s__:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus.s__) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Energie (kcal)`:`Energie, Gesamt (kcal)`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L7[is.na(cor.test.biom.L7)] <- 0


p.adj.biom.L7 <- subset(cor.test.biom.L7,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L7 <- subset(cor.test.biom.L7,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L7 <- log2(r.cor.biom.L7) %>% as.numeric(.)
cor.test.log.biom.L7[is.nan(cor.test.log.biom.L7)] <- 0
cor.test.log.biom.L7[is.infinite(cor.test.log.biom.L7)] <- 0
cor.test.log.biom.L7 <- as.data.frame(cor.test.log.biom.L7)

r.cor.biom.L7[is.nan(r.cor.biom.L7)] <- 0
r.cor.biom.L7[is.infinite(r.cor.biom.L7)] <- 0
r.cor.biom.L7 <- as.data.frame(r.cor.biom.L7)

#y axis -log10 p.value

p.adj.log.biom.L7 <- -log10(p.adj.biom.L7)
p.adj.log.biom.L7[is.infinite(p.adj.log.biom.L7)] <- 0
p.adj.log.biom.L7 <- as.data.frame(p.adj.log.biom.L7)

cor.test.adj.biom.L7 <- cbind(cor.test.biom.L7[ ,c(1,3)],p.adj.log.biom.L7) %>% cbind(.,r.cor.biom.L7) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L7) <- c('specy','nutrient','p.adj.log','estimate')

cor.test.adj.biom.L7 <- cor.test.adj.biom.L7[cor.test.adj.biom.L7$p.adj.log != 0 & cor.test.adj.biom.L7$estimate != 0,]

write_tsv(cor.test.adj.biom.L7,paste0(table_dir,'corr.nutr.biom.L7.tsv'))

```

###U1 correlation between nutrients and biom.L7.U1
```{r}

merge.biom.L7.nutr.U1 <- biom.L7.rel.clr.U1 %>% merge(.,nutr.sum.z.U1,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L7.nutr.U1[is.na(merge.biom.L7.nutr.U1)] <- 0

merge.biom.L7.nutr.U1$file_name <- row.names(merge.biom.L7.nutr.U1)
write_tsv(merge.biom.L7.nutr.U1,paste0(table_dir,'corr.biom.L7.U1.19.02.tsv',append=FALSE))
merge.biom.L7.nutr.U1 <- select(merge.biom.L7.nutr.U1,everything(),-starts_with('file_name'))

cor.test.biom.L7.U1 <- merge.biom.L7.nutr.U1 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter.s__:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus.s__) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Energie (kcal)`:`Energie, Gesamt (kcal)`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L7.U1[is.na(cor.test.biom.L7.U1)] <- 0


p.adj.biom.L7.U1 <- subset(cor.test.biom.L7.U1,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L7.U1 <- subset(cor.test.biom.L7.U1,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L7.U1 <- log2(r.cor.biom.L7.U1) %>% as.numeric(.)
cor.test.log.biom.L7.U1[is.nan(cor.test.log.biom.L7.U1)] <- 0
cor.test.log.biom.L7.U1[is.infinite(cor.test.log.biom.L7.U1)] <- 0
cor.test.log.biom.L7.U1 <- as.data.frame(cor.test.log.biom.L7.U1)

r.cor.biom.L7.U1[is.nan(r.cor.biom.L7.U1)] <- 0
r.cor.biom.L7.U1[is.infinite(r.cor.biom.L7.U1)] <- 0
r.cor.biom.L7.U1 <- as.data.frame(r.cor.biom.L7.U1)

#y axis -log10 p.value

p.adj.log.biom.L7.U1 <- -log10(p.adj.biom.L7.U1)
p.adj.log.biom.L7.U1[is.infinite(p.adj.log.biom.L7.U1)] <- 0
p.adj.log.biom.L7.U1 <- as.data.frame(p.adj.log.biom.L7.U1)

cor.test.adj.biom.L7.U1 <- cbind(cor.test.biom.L7.U1[ ,c(1,3)],p.adj.log.biom.L7.U1) %>% cbind(.,r.cor.biom.L7.U1) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L7.U1) <- c('genus','nutrient','p.adj.log','estimate')

cor.test.adj.biom.L7.U1 <- cor.test.adj.biom.L7.U1[cor.test.adj.biom.L7.U1$p.adj.log != 0 & cor.test.adj.biom.L7.U1$estimate != 0,]

write_tsv(cor.test.adj.biom.L7.U1,paste0(table_dir,'corr.nutr.biom.L7.U1.tsv'))

```

###U2 correlation between nutrients and biom.L7.U2
```{r}

merge.biom.L7.nutr.U2 <- biom.L7.rel.clr.U2 %>% merge(.,nutr.sum.z.U2,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L7.nutr.U2[is.na(merge.biom.L7.nutr.U2)] <- 0

merge.biom.L7.nutr.U2$file_name <- row.names(merge.biom.L7.nutr.U2)
write_tsv(merge.biom.L7.nutr.U2,paste0(table_dir,'corr.biom.L7.U1.19.02.tsv',append=FALSE))
merge.biom.L7.nutr.U2 <- select(merge.biom.L7.nutr.U2,everything(),-starts_with('file_name'))

cor.test.biom.L7.U2 <- merge.biom.L7.nutr.U2 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter.s__:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus.s__) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Energie (kcal)`:`Energie, Gesamt (kcal)`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L7.U2[is.na(cor.test.biom.L7.U2)] <- 0


p.adj.biom.L7.U2 <- subset(cor.test.biom.L7.U2,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L7.U2 <- subset(cor.test.biom.L7.U2,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L7.U2 <- log2(r.cor.biom.L7.U2) %>% as.numeric(.)
cor.test.log.biom.L7.U2[is.nan(cor.test.log.biom.L7.U2)] <- 0
cor.test.log.biom.L7.U2[is.infinite(cor.test.log.biom.L7.U2)] <- 0
cor.test.log.biom.L7.U2 <- as.data.frame(cor.test.log.biom.L7.U2)

r.cor.biom.L7.U2[is.nan(r.cor.biom.L7.U2)] <- 0
r.cor.biom.L7.U2[is.infinite(r.cor.biom.L7.U2)] <- 0
r.cor.biom.L7.U2 <- as.data.frame(r.cor.biom.L7.U2)

#y axis -log10 p.value

p.adj.log.biom.L7.U2 <- -log10(p.adj.biom.L7.U2)
p.adj.log.biom.L7.U2[is.infinite(p.adj.log.biom.L7.U2)] <- 0
p.adj.log.biom.L7.U2 <- as.data.frame(p.adj.log.biom.L7.U2)

cor.test.adj.biom.L7.U2 <- cbind(cor.test.biom.L7.U2[ ,c(1,3)],p.adj.log.biom.L7.U2) %>% cbind(.,r.cor.biom.L7.U2) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L7.U2) <- c('specy','nutrient','p.adj.log','estimate')

cor.test.adj.biom.L7.U2 <- cor.test.adj.biom.L7.U2[cor.test.adj.biom.L7.U2$p.adj.log != 0 & cor.test.adj.biom.L7.U2$estimate != 0,]

write_tsv(cor.test.adj.biom.L7.U2,paste0(table_dir,'corr.nutr.biom.L7.U2.tsv'))
```

#volcano plots
```{r}
##sig. correlations
sig.biom.L6 <- cor.test.adj.biom.L6[cor.test.adj.biom.L6$p.adj.log >= -log10(0.050) & cor.test.adj.biom.L6$p.adj.log > 0.00000,]
top.sig.biom.L6 <- sig.biom.L6[order(-sig.biom.L6$p.adj.log ),] #[1:11,]

sig.biom.L7 <- cor.test.adj.biom.L7[cor.test.adj.biom.L7$p.adj.log >= -log10(0.050) & cor.test.adj.biom.L7$p.adj.log > 0.00000,]
top.sig.biom.L7 <- sig.biom.L7[order(-sig.biom.L7$p.adj.log ),] #[1:11,]

##L6
vp.biom.L6 <- ggplot(cor.test.adj.biom.L6,
    aes(x = `estimate`, y = `p.adj.log`)) +
  geom_point() +
  ggtitle("Volcano plot") +
  xlab("correlation") + 
  ylab("-log10 adjusted p-value")

vp.biom.L6 <- vp.biom.L6+xlim(c(-1,1)) + geom_vline(xintercept=c(-0.6, 0.6), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red") 

##L7
vp.biom.L7 <- ggplot(cor.test.adj.biom.L7,
    aes(x = `estimate`, y = `p.adj.log`)) +
  geom_point() +
  ggtitle("Volcano plot") +
  xlab("correlation") + 
  ylab("-log10 adjusted p-value")

vp.biom.L7 <- vp.biom.L7+xlim(c(-1,1)) + geom_vline(xintercept=c(-0.6, 0.6), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red")
```