---
title: "correlation rarefied taxa with conventional clusters (all,U1,U2)"
output: html_notebook
---

#libraries
```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("biomformat")

# if (!requireNamespace('BiocManager', quietly = TRUE))
#     install.packages('BiocManager')
# 
#  BiocManager::install('EnhancedVolcano')

# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("a4Base")
  

library(tidyverse)
library(biomformat)
# library(data.table)
library(compositions)
# library(corrplot)
# library(EnhancedVolcano)
library(a4Base)
library(plotly)
library(ggpubr)

library(Maaslin2)
```

# Paths
```{r}
wd <- '../../' 

output_dir = paste0(wd,'results/')
diet.protocols = paste0(output_dir,'diet.protocols/')

data_dir = paste0(wd,'data/')
table_dir = paste0(data_dir, 'tables/')
biom_dir = paste0(data_dir,'silva.feces/')
```

## Create Output Dirs
```{r}
dir.create(output_dir)
dir.create(table_dir)
```

# Data Import
```{r}

norm.L6 <- read_tsv(paste0(table_dir,'biom.L6.rel.clr_19.02.tsv')) %>% column_to_rownames(.,'rowname')
norm.L7 <- read_tsv(paste0(table_dir,'biom.L7.rel.clr_19.02.tsv'))%>% column_to_rownames(.,'rowname')

norm.L6.U1 <- read_tsv(paste0(table_dir,'biom.L6.rel.clr.U1_19.02.tsv'))%>% column_to_rownames(.,'rowname')
norm.L7.U1 <-read_tsv(paste0(table_dir,'biom.L7.rel.clr.U1_19.02.tsv'))%>% column_to_rownames(.,'rowname')

norm.L6.U2 <-read_tsv(paste0(table_dir,'biom.L6.rel.clr.U2_19.02.tsv'))%>% column_to_rownames(.,'rowname')
norm.L7.U2 <-read_tsv(paste0(table_dir,'biom.L7.rel.clr.U2_19-02.tsv'))%>% column_to_rownames(.,'rowname')

conv.clust.z <- read_tsv(paste0(table_dir, 'conv.cluster.correlation_17.02.tsv'))%>% column_to_rownames(.,'file_name')

conv.clust.z.U1 <- conv.clust.z[grepl('U1',row.names(conv.clust.z)),]
conv.clust.z.U2 <- conv.clust.z[grepl('U2',row.names(conv.clust.z)),]

```

###biom.L6 and conventional cluster correlation

merge relative clr taxa table with standardized conv cluster table
correlate every conv cluster with every taxa
-log10(BH p.value)

conv.sum.z kommt von biom.microbiome
```{r}
merge.L6.nutr <- norm.L6 %>% merge(.,conv.clust.z,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.L6.nutr[is.na(merge.L6.nutr)] <- 0


cor.test.L6.nutr <- merge.L6.nutr %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Alkoholfreie Getränke`:`Wurst, Fleischwaren`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.L6.nutr[is.na(cor.test.L6.nutr)] <- 0

p.adj.L6.nutr <- subset(cor.test.L6.nutr,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.L6.nutr <- subset(cor.test.L6.nutr,select = `estimate`) %>% data.matrix(.)
```

```{r}
#x axis log2 correlation

# cor.test.log.L6 <- log2(r.cor.L6) %>% as.numeric(.)
# cor.test.log.L6[is.nan(cor.test.log.L6)] <- 0
# cor.test.log.L6[is.infinite(cor.test.log.L6)] <- 0
# cor.test.log.L6 <- as.data.frame(cor.test.log.L6)

r.cor.L6.nutr[is.nan(r.cor.L6.nutr)] <- 0
r.cor.L6.nutr[is.infinite(r.cor.L6.nutr)] <- 0
r.cor.L6.nutr <- as.data.frame(r.cor.L6.nutr)

#y axis -log10 p.value

p.adj.log.L6.nutr <- -log10(p.adj.L6.nutr)
p.adj.log.L6.nutr[is.infinite(p.adj.log.L6.nutr)] <- 0
p.adj.log.L6.nutr <- as.data.frame(p.adj.log.L6.nutr)

cor.test.adj.L6.nutr <- cbind(cor.test.L6.nutr[ ,c(1,3)],p.adj.log.L6.nutr) %>% cbind(.,r.cor.L6.nutr) %>% as.data.frame(.)
colnames(cor.test.adj.L6.nutr) <- c('genus','clusters','p.adj.log','estimate')

cor.test.adj.L6.nutr <- cor.test.adj.L6.nutr[cor.test.adj.L6.nutr$p.adj.log != 0 & cor.test.adj.L6.nutr$estimate != 0,]

write_tsv(cor.test.adj.L6.nutr,paste0(table_dir,'corr.convclust.biom.L6.tsv'))

sig.nutr.L6 <- cor.test.adj.L6.nutr[cor.test.adj.L6.nutr$p.adj.log >= -log10(0.05) & cor.test.adj.L6.nutr$p.adj.log > 0.00000  ,] 
top.sig.nutr.L6 <- sig.nutr[order(-sig.nutr$p.adj.log ),]

```

###biom.L7 and convclust correlatin
```{r}
merge.L7.nutr <- norm.L7 %>% merge(.,conv.clust.z,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.L7.nutr[is.na(merge.L7.nutr)] <- 0


cor.test.L7.nutr <- merge.L7.nutr %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter.s__:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus.s__) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Alkoholfreie Getränke`:`Wurst, Fleischwaren`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.L7.nutr[is.na(cor.test.L7.nutr)] <- 0

p.adj.L7.nutr <- subset(cor.test.L7.nutr,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.L7.nutr <- subset(cor.test.L7.nutr,select = `estimate`) %>% data.matrix(.)
```

```{r}
#x axis log2 correlation

# cor.test.log.L7 <- log2(r.cor.L7) %>% as.numeric(.)
# cor.test.log.L7[is.nan(cor.test.log.L7)] <- 0
# cor.test.log.L7[is.infinite(cor.test.log.L7)] <- 0
# cor.test.log.L7 <- as.data.frame(cor.test.log.L7)

r.cor.L7.nutr[is.nan(r.cor.L7.nutr)] <- 0
r.cor.L7.nutr[is.infinite(r.cor.L7.nutr)] <- 0
r.cor.L7.nutr <- as.data.frame(r.cor.L7.nutr)

#y axis -log10 p.value

p.adj.log.L7.nutr <- -log10(p.adj.L7.nutr)
p.adj.log.L7.nutr[is.infinite(p.adj.log.L7.nutr)] <- 0
p.adj.log.L7.nutr <- as.data.frame(p.adj.log.L7.nutr)

cor.test.adj.L7.nutr <- cbind(cor.test.L7.nutr[ ,c(1,3)],p.adj.log.L7.nutr) %>% cbind(.,r.cor.L7.nutr) %>% as.data.frame(.)
colnames(cor.test.adj.L7.nutr) <- c('specy','clusters','p.adj.log','estimate')

cor.test.adj.L7.nutr <- cor.test.adj.L7.nutr[cor.test.adj.L7.nutr$p.adj.log != 0 & cor.test.adj.L7.nutr$estimate != 0,]

write_tsv(cor.test.adj.L7.nutr,paste0(table_dir,'corr.convclust.biom.L7.tsv'))

sig.nutr.L7 <- cor.test.adj.L7.nutr[cor.test.adj.L7.nutr$p.adj.log >= -log10(0.05) & cor.test.adj.L7.nutr$p.adj.log > 0.00000  ,] 
top.sig.nutr.L7 <- sig.nutr[order(-sig.nutr$p.adj.log ),]

```

## L6 U1 and U2 biom.L6 and conv clusts
###U1
```{r}

merge.biom.L6.nutr.U1 <- norm.L6.U1 %>% merge(.,conv.clust.z.U1,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L6.nutr.U1[is.na(merge.biom.L6.nutr.U1)] <- 0

merge.biom.L6.nutr.U1$file_name <- row.names(merge.biom.L6.nutr.U1)
write_tsv(merge.biom.L6.nutr.U1,'C:/Users/Juliane/Desktop/project.directory/data/tables/corr.biom.L6.convclust.U1.19.02.tsv',append=FALSE)
merge.biom.L6.nutr.U1 <- select(merge.biom.L6.nutr.U1,everything(),-starts_with('file_name'))

cor.test.biom.L6.U1 <- merge.biom.L6.nutr.U1 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Alkoholfreie Getränke`:`Wurst, Fleischwaren`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L6.U1[is.na(cor.test.biom.L6.U1)] <- 0


p.adj.biom.L6.U1 <- subset(cor.test.biom.L6.U1,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L6.U1 <- subset(cor.test.biom.L6.U1,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L6.U1 <- log2(r.cor.biom.L6.U1) %>% as.numeric(.)
cor.test.log.biom.L6.U1[is.nan(cor.test.log.biom.L6.U1)] <- 0
cor.test.log.biom.L6.U1[is.infinite(cor.test.log.biom.L6.U1)] <- 0
cor.test.log.biom.L6.U1 <- as.data.frame(cor.test.log.biom.L6.U1)

r.cor.biom.L6.U1[is.nan(r.cor.biom.L6.U1)] <- 0
r.cor.biom.L6.U1[is.infinite(r.cor.biom.L6.U1)] <- 0
r.cor.biom.L6.U1 <- as.data.frame(r.cor.biom.L6.U1)

#y axis -log10 p.value

p.adj.log.biom.L6.U1 <- -log10(p.adj.biom.L6.U1)
p.adj.log.biom.L6.U1[is.infinite(p.adj.log.biom.L6.U1)] <- 0
p.adj.log.biom.L6.U1 <- as.data.frame(p.adj.log.biom.L6.U1)

cor.test.adj.biom.L6.U1 <- cbind(cor.test.biom.L6.U1[ ,c(1,3)],p.adj.log.biom.L6.U1) %>% cbind(.,r.cor.biom.L6.U1) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L6.U1) <- c('genus','clusters','p.adj.log','estimate')

cor.test.adj.biom.L6.U1 <- cor.test.adj.biom.L6.U1[cor.test.adj.biom.L6.U1$p.adj.log != 0 & cor.test.adj.biom.L6.U1$estimate != 0,]

write_tsv(cor.test.adj.biom.L6.U1,paste0(table_dir,'corr.convclust.biom.L6.U1.tsv'))

```

###U2 
```{r}

merge.biom.L6.nutr.U2 <- norm.L6.U2 %>% merge(.,conv.clust.z.U2,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L6.nutr.U2[is.na(merge.biom.L6.nutr.U2)] <- 0

merge.biom.L6.nutr.U2$file_name <- row.names(merge.biom.L6.nutr.U2)
write_tsv(merge.biom.L6.nutr.U2,paste0(table_dir,'corr.biom.L6.U2.convclust.19.02.tsv',append=FALSE))
merge.biom.L6.nutr.U2 <- select(merge.biom.L6.nutr.U2,everything(),-starts_with('file_name'))

cor.test.biom.L6.U2 <- merge.biom.L6.nutr.U2 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Alkoholfreie Getränke`:`Wurst, Fleischwaren`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L6.U2[is.na(cor.test.biom.L6.U2)] <- 0


p.adj.biom.L6.U2 <- subset(cor.test.biom.L6.U2,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L6.U2 <- subset(cor.test.biom.L6.U2,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L6.U2 <- log2(r.cor.biom.L6.U2) %>% as.numeric(.)
cor.test.log.biom.L6.U2[is.nan(cor.test.log.biom.L6.U2)] <- 0
cor.test.log.biom.L6.U2[is.infinite(cor.test.log.biom.L6.U2)] <- 0
cor.test.log.biom.L6.U2 <- as.data.frame(cor.test.log.biom.L6.U2)

r.cor.biom.L6.U2[is.nan(r.cor.biom.L6.U2)] <- 0
r.cor.biom.L6.U2[is.infinite(r.cor.biom.L6.U2)] <- 0
r.cor.biom.L6.U2 <- as.data.frame(r.cor.biom.L6.U2)

#y axis -log10 p.value

p.adj.log.biom.L6.U2 <- -log10(p.adj.biom.L6.U2)
p.adj.log.biom.L6.U2[is.infinite(p.adj.log.biom.L6.U2)] <- 0
p.adj.log.biom.L6.U2 <- as.data.frame(p.adj.log.biom.L6.U2)

cor.test.adj.biom.L6.U2 <- cbind(cor.test.biom.L6.U2[ ,c(1,3)],p.adj.log.biom.L6.U2) %>% cbind(.,r.cor.biom.L6.U2) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L6.U2) <- c('genus','clusters','p.adj.log','estimate')

cor.test.adj.biom.L6.U2 <- cor.test.adj.biom.L6.U2[cor.test.adj.biom.L6.U2$p.adj.log != 0 & cor.test.adj.biom.L6.U2$estimate != 0,]

write_tsv(cor.test.adj.biom.L6.U2,paste0(table_dir,'corr.convclust.biom.L6.U2.tsv'))

```

## L7 U1 and U2 biom.L7 and conv clusts
###U1 
```{r}

merge.biom.L7.nutr.U1 <- biom.L7.rel.clr.U1 %>% merge(.,conv.clust.z.U1,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L7.nutr.U1[is.na(merge.biom.L7.nutr.U1)] <- 0

merge.biom.L7.nutr.U1$file_name <- row.names(merge.biom.L7.nutr.U1)
write_tsv(merge.biom.L7.nutr.U1,paste0(table_dir,'corr.biom.L7.U1.nutclust.19.02.tsv',append=FALSE))
merge.biom.L7.nutr.U1 <- select(merge.biom.L7.nutr.U1,everything(),-starts_with('file_name'))

cor.test.biom.L7.U1 <- merge.biom.L7.nutr.U1 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter.s__:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus.s__) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Alkoholfreie Getränke`:`Wurst, Fleischwaren`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L7.U1[is.na(cor.test.biom.L7.U1)] <- 0


p.adj.biom.L7.U1 <- subset(cor.test.biom.L7.U1,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L7.U1 <- subset(cor.test.biom.L7.U1,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L7.U1 <- log2(r.cor.biom.L7.U1) %>% as.numeric(.)
cor.test.log.biom.L7.U1[is.nan(cor.test.log.biom.L7.U1)] <- 0
cor.test.log.biom.L7.U1[is.infinite(cor.test.log.biom.L7.U1)] <- 0
cor.test.log.biom.L7.U1 <- as.data.frame(cor.test.log.biom.L7.U1)

r.cor.biom.L7.U1[is.nan(r.cor.biom.L7.U1)] <- 0
r.cor.biom.L7.U1[is.infinite(r.cor.biom.L7.U1)] <- 0
r.cor.biom.L7.U1 <- as.data.frame(r.cor.biom.L7.U1)

#y axis -log10 p.value

p.adj.log.biom.L7.U1 <- -log10(p.adj.biom.L7.U1)
p.adj.log.biom.L7.U1[is.infinite(p.adj.log.biom.L7.U1)] <- 0
p.adj.log.biom.L7.U1 <- as.data.frame(p.adj.log.biom.L7.U1)

cor.test.adj.biom.L7.U1 <- cbind(cor.test.biom.L7.U1[ ,c(1,3)],p.adj.log.biom.L7.U1) %>% cbind(.,r.cor.biom.L7.U1) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L7.U1) <- c('specy','clusters','p.adj.log','estimate')

cor.test.adj.biom.L7.U1 <- cor.test.adj.biom.L7.U1[cor.test.adj.biom.L7.U1$p.adj.log != 0 & cor.test.adj.biom.L7.U1$estimate != 0,]

write_tsv(cor.test.adj.biom.L7.U1,paste0(table_dir,'corr.convclust.biom.L7.U1.tsv'))

```

###U2 
```{r}

merge.biom.L7.nutr.U2 <- biom.L7.rel.clr.U2 %>% merge(.,conv.clust.z.U2,by = 'row.names',all.x=T,all.y = T) %>% column_to_rownames(.,'Row.names')

merge.biom.L7.nutr.U2[is.na(merge.biom.L7.nutr.U2)] <- 0

merge.biom.L7.nutr.U2$file_name <- row.names(merge.biom.L7.nutr.U2)
write_tsv(merge.biom.L7.nutr.U2,paste0(table_dir,'corr.biom.L7.U1.convclust.19.02.tsv',append=FALSE))
merge.biom.L7.nutr.U2 <- select(merge.biom.L7.nutr.U2,everything(),-starts_with('file_name'))

cor.test.biom.L7.U2 <- merge.biom.L7.nutr.U2 %>% 
  pivot_longer(names_to = "variable_right", values_to = "value_right", k__Archaea.p__Euryarchaeota.c__Methanobacteria.o__Methanobacteriales.f__Methanobacteriaceae.g__Methanobrevibacter.s__:k__Bacteria.p__.Thermi..c__Deinococci.o__Thermales.f__Thermaceae.g__Meiothermus.s__) %>% 
  pivot_longer(names_to = "variable_left", values_to = "value_left", `Alkoholfreie Getränke`:`Wurst, Fleischwaren`) %>% 
  group_by(variable_left, variable_right) %>% 
  
  mutate(p.value = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$p.value,
         estimate = broom::tidy(cor.test(value_left, value_right,method = 'spearman'))$estimate)

cor.test.biom.L7.U2[is.na(cor.test.biom.L7.U2)] <- 0


p.adj.biom.L7.U2 <- subset(cor.test.biom.L7.U2,select = `p.value`) %>% data.matrix(.) %>% p.adjust(.,method = 'fdr')
r.cor.biom.L7.U2 <- subset(cor.test.biom.L7.U2,select = `estimate`) %>% data.matrix(.)

```

plotting
```{r}

#x axis log2 correlation

cor.test.log.biom.L7.U2 <- log2(r.cor.biom.L7.U2) %>% as.numeric(.)
cor.test.log.biom.L7.U2[is.nan(cor.test.log.biom.L7.U2)] <- 0
cor.test.log.biom.L7.U2[is.infinite(cor.test.log.biom.L7.U2)] <- 0
cor.test.log.biom.L7.U2 <- as.data.frame(cor.test.log.biom.L7.U2)

r.cor.biom.L7.U2[is.nan(r.cor.biom.L7.U2)] <- 0
r.cor.biom.L7.U2[is.infinite(r.cor.biom.L7.U2)] <- 0
r.cor.biom.L7.U2 <- as.data.frame(r.cor.biom.L7.U2)

#y axis -log10 p.value

p.adj.log.biom.L7.U2 <- -log10(p.adj.biom.L7.U2)
p.adj.log.biom.L7.U2[is.infinite(p.adj.log.biom.L7.U2)] <- 0
p.adj.log.biom.L7.U2 <- as.data.frame(p.adj.log.biom.L7.U2)

cor.test.adj.biom.L7.U2 <- cbind(cor.test.biom.L7.U2[ ,c(1,3)],p.adj.log.biom.L7.U2) %>% cbind(.,r.cor.biom.L7.U2) %>% as.data.frame(.)
colnames(cor.test.adj.biom.L7.U2) <- c('specy','clusters','p.adj.log','estimate')

cor.test.adj.biom.L7.U2 <- cor.test.adj.biom.L7.U2[cor.test.adj.biom.L7.U2$p.adj.log != 0 & cor.test.adj.biom.L7.U2$estimate != 0,]

write_tsv(cor.test.adj.biom.L7.U2,paste0(table_dir,'corr.convclust.biom.L7.U2.tsv'))
```